# Get Started {#sec-start}

```{r setup-start, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  message = FALSE,
  warning = FALSE
)
```

This chapter covers only the setup needed to run the tutorial: install or load
`fdid`, load the example data, and prepare the wide panel shape used by
`fdid()`.

## Installation

Install the package from GitHub for ordinary use:

```{r install-github, eval=FALSE}
install.packages("devtools")
devtools::install_github("xuyiqing/fdid")
```

When reviewing a local source checkout, install the local source into the R
library that will be used for rendering:

```{bash install-local, eval=FALSE}
R CMD build fdid
R CMD INSTALL fdid_<version>.tar.gz
```

After installation, restart R and confirm the package path:

```{r load-and-check, eval=FALSE}
library(fdid)
packageVersion("fdid")
normalizePath(find.package("fdid"))
```

## Dataset

The tutorial uses the package's famine mortality data. The raw data are long
panel data: each unit has multiple year rows, an outcome, a baseline factor
`G`, and covariates. `fdid_prepare()` converts this layout to one row per unit
with one outcome column per time period.

```{r data-shape, eval=FALSE}
library(fdid)
data(fdid)

mortality$uniqueid <- paste(
  mortality$provid,
  mortality$countyid,
  sep = "-"
)

covar <- c("avggrain", "nograin", "urban", "dis_bj", "dis_pc",
           "rice", "minority", "edu", "lnpop")

s <- fdid_prepare(
  data = mortality,
  Y_label = "mortality",
  X_labels = covar,
  G_label = "pczupu",
  unit_label = "uniqueid",
  time_label = "year"
)

str(s)
```

The prepared object is the input to `fdid()`. It contains `G`, covariates, and
wide outcome columns such as `Y_1957`, `Y_1958`, and later years. The following
chapters reuse this pattern for binary, discrete, and continuous baseline
factors.
