# Binary G {#sec-binary}

```{r setup-binary, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  message = FALSE,
  warning = FALSE
)
```

This chapter shows the original binary/discrete-`G` FDID workflow. Use this
path when the estimand is a treated-versus-control contrast or a comparison
between a small number of baseline-factor groups.

## Data preparation

`fdid_prepare()` expects long panel data and returns the wide object consumed
by `fdid()`.

```{r binary-prepare, eval=FALSE}
library(fdid)
data(fdid)

mortality$uniqueid <- paste(mortality$provid, mortality$countyid, sep = "-")
mortality$highzupu <- ifelse(
  mortality$pczupu >= median(mortality$pczupu, na.rm = TRUE),
  1,
  0
)

covar <- c("avggrain", "nograin", "urban", "dis_bj", "dis_pc",
           "rice", "minority", "edu", "lnpop")

s_binary <- fdid_prepare(
  data = mortality,
  Y_label = "mortality",
  X_labels = covar,
  G_label = "highzupu",
  unit_label = "uniqueid",
  time_label = "year"
)
```

If cluster-robust inference is needed, pass `cluster_label` to
`fdid_prepare()` so the prepared object carries the clustering variable through
estimation.

## FDID: A quick review

FDID compares how outcomes change around a common event across units with
different baseline factor values. For the famine example, the event period is
1958-1961 and the reference year is 1957.

```{r binary-periods, eval=FALSE}
tr_period <- 1958:1961
ref_period <- 1957
```

## Estimation

The binary/discrete workflow uses the traditional methods exposed through the
same `fdid()` interface.

### Difference-in-Differences

```{r did-fit, eval=FALSE}
fit_did <- fdid(
  s_binary,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "did",
  vartype = "robust"
)
summary(fit_did)
```

### OLS without interactions

```{r ols1-fit, eval=FALSE}
fit_ols1 <- fdid(
  s_binary,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "ols1",
  vartype = "robust"
)
summary(fit_ols1)
```

### OLS with interactions

```{r ols2-fit, eval=FALSE}
fit_ols2 <- fdid(
  s_binary,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "ols2",
  vartype = "robust"
)
summary(fit_ols2)
```

### Entropy balancing

```{r ebal-fit, eval=FALSE}
fit_ebal <- fdid(
  s_binary,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "ebal",
  target.pop = "1",
  vartype = "robust"
)
summary(fit_ebal)
```

### IPW

```{r ipw-fit, eval=FALSE}
fit_ipw <- fdid(
  s_binary,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "ipw",
  vartype = "robust"
)
summary(fit_ipw)
```

### AIPW

```{r aipw-fit, eval=FALSE}
fit_aipw <- fdid(
  s_binary,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "aipw",
  vartype = "robust"
)
summary(fit_aipw)
```

## Inference

Binary/discrete methods support robust, bootstrap, and jackknife variance
choices where available. Use robust inference for fast first checks and
bootstrap or jackknife when the design calls for a resampling comparison.

```{r binary-inference, eval=FALSE}
fit_ols1_boot <- fdid(
  s_binary,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "ols1",
  vartype = "bootstrap",
  nsims = 199,
  parallel = FALSE
)
```

## Estimation options

`target.pop` controls whether weighted estimators target the `G = 1`,
`G = 0`, or combined population where the method supports that choice.
`entire_period` controls whether all periods or a shorter event window enter
the dynamic output. `missing_data` controls how incomplete rows are handled.

## Plotting the results

### Raw means by group

```{r plot-raw, eval=FALSE}
plot(fit_ols1, type = "raw", connected = TRUE)
```

### Dynamic estimates

```{r plot-dynamic, eval=FALSE}
plot(fit_ols1, type = "dynamic", connected = TRUE)
```

### Diagnostic test for overlap

```{r plot-overlap, eval=FALSE}
plot(fit_aipw, type = "overlap")
```

### Comparison of methods

```{r plot-list, eval=FALSE}
fits <- fdid_list(fit_did, fit_ols1, fit_ols2, fit_ebal, fit_ipw, fit_aipw)
plot(fits)
```

The continuous-`G` chapters start from a different question. Once `G` is
continuous, a scalar slope, level curve, derivative curve, average derivative,
and fixed-reference contrast are different targets. Choose the target first,
then choose the method.
