# Continuous-G Overview {#sec-continuous-overview}

```{r setup-continuous-overview, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  message = FALSE,
  warning = FALSE
)
```

This chapter is the target map for continuous-`G` analysis. It does not try to
fit every estimator. It explains which object you want, how to prepare the
design, why support comes before estimation, and which chapter gives the full
workflow.

## Choose the Target First

Continuous-`G` methods do not estimate one interchangeable object. Decide
whether the report needs a scalar, a curve, a derivative, a contrast, or a
diagnostic display before choosing `method`.

| User target | Primary workflow | Reporting/display |
|---|---|---|
| Binary-G or discrete treated-vs-control FDID effect | `method = "did"`, `"ols1"`, `"ols2"`, `"ebal"`, `"ipw"`, `"aipw"` | `summary()`, raw/dynamic/overlap plots |
| Continuous-G scalar linear slope | `method = "dml_plr"` | `summary()` with scalar DML inference label |
| Smooth continuous-G level curve | `method = "kernel"` first; `method = "dml_flex"` for adjusted flexible nuisance learning | `plot(type = "curve", curve = "level")`, curve table |
| Continuous-G derivative curve | `kernel` or `dml_flex` with derivative output | `plot(type = "curve", curve = "derivative")`, `fdid_derivative()` |
| Observed-population average derivative | `method = "dml_incremental"` | `summary()` and density diagnostics |
| Fixed-reference level contrast | fitted `kernel` or `dml_flex` plus `fdid_contrast()` | `fdid_contrast()`, `plot(type = "contrast")` |
| Support and overlap diagnostics | preparation and plotting diagnostics | support panels, rugs, density/histogram displays, overlap plots |

Keep four concepts separate:

1. the estimand, such as a scalar slope or `theta(g1) - theta(g0)`;
2. the estimator, such as `kernel`, `dml_flex`, or `dml_incremental`;
3. the inference calculation, such as analytical, bootstrap, or multiplier;
4. the plot or reporting view, such as a curve plot or contrast table.

## Prepare the Continuous-G Design

Use a continuous baseline factor in `fdid_prepare()`. The example below uses
`lnpczupu`, a log social-capital measure from the mortality data.

```{r continuous-prepare, eval=FALSE}
library(fdid)
data(fdid)

mortality$uniqueid <- paste(mortality$provid, mortality$countyid, sep = "-")

covar <- c("avggrain", "nograin", "urban", "dis_bj", "dis_pc",
           "rice", "minority", "edu", "lnpop")

s_cont <- fdid_prepare(
  data = transform(mortality, G_cont = lnpczupu),
  Y_label = "mortality",
  X_labels = covar,
  G_label = "G_cont",
  unit_label = "uniqueid",
  time_label = "year"
)

tr_period <- 1958:1961
ref_period <- 1957
```

For curve methods, `eval_g` is an estimation grid. The package estimates and
stores curve quantities at those values; plotting displays the stored grid
rather than densifying the curve after the fact.

```{r continuous-grid, eval=FALSE}
g_positive <- s_cont$G[s_cont$G > 0]
eval_g <- as.numeric(stats::quantile(
  g_positive,
  probs = seq(0.20, 0.80, length.out = 9),
  na.rm = TRUE
))
```

## Inspect Support Before Estimation

Continuous-`G` curve estimates are weakest where observed support is sparse.
Inspect the empirical distribution of `G` before fitting a curve and avoid
turning a tail extrapolation into a headline result.

```{r continuous-support, eval=FALSE}
hist(s_cont$G, breaks = 35, col = "gray85", border = "white",
     xlab = "Baseline factor G", main = "Observed support")
abline(v = range(eval_g), col = c("steelblue", "firebrick"), lwd = 2)
rug(s_cont$G, col = grDevices::adjustcolor("black", 0.25))
```

## Method Choice In One Table

| Need | Use first | Then check | Avoid saying |
|---|---|---|---|
| Quick scalar benchmark | `dml_plr` | OLS/discrete sensitivity | "This is the whole dose-response curve" |
| Smooth level curve | `kernel` | bandwidth and support | "Kernel removes all confounding concerns" |
| Flexible adjusted level curve | `dml_flex` | learners, density, signal map | "The scalar summary is the main result" |
| Observed average derivative | `dml_incremental` | density diagnostics and fold sensitivity | "It is just the average plotted slope" |
| Fixed-reference comparison | `fdid_contrast()` after `kernel` or `dml_flex` | exact grid vs interpolation and available covariance/replicates | "All contrast intervals use the same variance source" |
| Plot support | curve plot support options | mass points and sparse regions | "The support display is adjusted causal evidence" |

## What Not To Compare Directly

Do not line up `dml_plr`, `kernel`, `dml_flex`, and `dml_incremental` as if
they estimate the same number. `dml_plr` is a scalar partially linear slope.
`kernel` and `dml_flex` can report level and derivative curves. `dml_incremental`
is an observed-population average derivative, not the average of a plotted
`dml_flex` derivative curve. A fixed-reference contrast is a function of a
stored level curve, not a new estimator.
