# Kernel Estimator for Continuous G {#sec-kernel-continuous}

```{r kernel-setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 7,
  fig.height = 4.8
)
set.seed(20260531)

fit_fdid <- function(...) {
  tryCatch(
    fdid(...),
    error = function(e) {
      structure(list(message = conditionMessage(e)), class = "fdid_fit_error")
    }
  )
}

is_fit_error <- function(x) inherits(x, "fdid_fit_error")

event_table <- function(obj, label) {
  if (is_fit_error(obj)) {
    return(data.frame(
      fit = label, Estimate = NA_real_, Std.Error = NA_real_,
      CI_Lower = NA_real_, CI_Upper = NA_real_, note = obj$message,
      check.names = FALSE
    ))
  }
  out <- obj$est$event
  data.frame(
    fit = label,
    Estimate = as.numeric(out$Estimate[1]),
    Std.Error = as.numeric(out$Std.Error[1]),
    CI_Lower = as.numeric(out$CI_Lower[1]),
    CI_Upper = as.numeric(out$CI_Upper[1]),
    note = "",
    check.names = FALSE
  )
}

kernel_curve_table <- function(obj, n = 8) {
  if (is_fit_error(obj) || is.null(obj$curve_event)) return(data.frame())
  ce <- obj$curve_event
  head(data.frame(
    G = obj$eval_g,
    mu_hat = ce$mu_hat,
    se_mu = ce$se_mu,
    delta_hat = ce$delta_hat,
    se_delta = ce$se_delta
  ), n)
}
```

This chapter gives a focused workflow for `method = "kernel"` with continuous
baseline factor `G`. The structure follows the practical style used in the
`interflex` tutorial and in @liu2025practical: diagnose support first, estimate
the curve, inspect uncertainty, then report specific contrasts or derivatives.

The kernel estimator is useful when the main empirical object is a smooth
continuous-$G$ curve rather than a single slope. It estimates two event-period
objects:

- the level curve, $\mu(g)$, the local FDID response at a value of `G`;
- the derivative curve, $\delta(g)$, the local slope of that response.

Unlike the DML estimators in Chapter @sec-dml-continuous, the kernel estimator
is a fused local smoother: the local linear fit, support weighting, and curve
aggregation are part of one estimator.

## Data Setup And Support

```{r kernel-library-data}
library(fdid)
data(fdid)

mortality$uniqueid <- paste(
  as.character(mortality$provid),
  as.character(mortality$countyid),
  sep = "-"
)

covar <- c("avggrain", "nograin", "urban", "dis_bj", "dis_pc",
           "rice", "minority", "edu", "lnpop")
tr_period <- 1958:1961
ref_period <- 1957

s_cont <- fdid_prepare(
  data = transform(mortality, G_cont = lnpczupu),
  Y_label = "mortality",
  X_labels = covar,
  G_label = "G_cont",
  unit_label = "uniqueid",
  time_label = "year"
)
```

The famine mortality example has a mass point at zero in `lnpczupu`. The grid
below keeps that lowest observed value as a reporting anchor, then adds
positive-support landmarks where the smooth curve is more interpretable.

```{r kernel-support-grid}
g_positive <- s_cont$G[s_cont$G > 0]

eval_g_kernel <- c(0, as.numeric(stats::quantile(
  g_positive,
  probs = c(0.20, 0.35, 0.50, 0.65, 0.80),
  na.rm = TRUE
)))

kernel_landmarks <- c(0, as.numeric(stats::quantile(
  g_positive,
  probs = c(0.20, 0.50, 0.80),
  na.rm = TRUE
)))

support_summary <- data.frame(
  quantity = c("n units", "share G = 0", "20% positive support",
               "80% positive support", "grid minimum", "grid maximum",
               "grid size"),
  value = c(nrow(s_cont), mean(s_cont$G == 0),
            min(eval_g_kernel[eval_g_kernel > 0]),
            max(eval_g_kernel), min(eval_g_kernel), max(eval_g_kernel),
            length(eval_g_kernel))
)

knitr::kable(support_summary, digits = 4)
```

```{r kernel-support-plot}
hist(s_cont$G, breaks = 35, col = "gray85", border = "white",
     xlab = "Baseline factor G: lnpczupu",
     main = "Observed support for continuous G")
abline(v = eval_g_kernel, col = "steelblue", lwd = 1.5)
abline(v = kernel_landmarks, col = "firebrick", lty = 3, lwd = 1.5)
rug(s_cont$G, col = grDevices::adjustcolor("black", 0.25))
legend("topright",
       legend = c("estimation grid", "reporting landmarks"),
       col = c("steelblue", "firebrick"), lty = c(1, 3), lwd = 1.5,
       bty = "n")
```

The `eval_g` grid is an estimation grid, not a plotting-only option. If five
values are supplied, `fdid()` estimates five local curve points. For publication
figures, use a denser grid over the region you intend to interpret.

## Basic Kernel Fit

The next fit uses a fixed bandwidth `h0` so that the tutorial renders quickly
and reproducibly. Leaving `h0 = NULL` asks the package to select the baseline
bandwidth by cross-validation.

```{r kernel-basic-fit}
fit_kernel_fast <- fit_fdid(
  s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "kernel",
  eval_g = eval_g_kernel,
  h0 = 0.18,
  K_folds = 3,
  vartype = "robust"
)

knitr::kable(event_table(fit_kernel_fast, "kernel robust"), digits = 3)
```

The scalar event row is a compact summary. For the kernel estimator, the main
objects are the stored curve arrays.

```{r kernel-curve-table}
if (!is_fit_error(fit_kernel_fast)) {
  knitr::kable(kernel_curve_table(fit_kernel_fast), digits = 3)
}
```

## Level And Derivative Curves

`plot(type = "curve")` draws the stored curve values. Use `curve = "level"` for
the FDID response curve and `curve = "derivative"` for the local slope.

```{r kernel-basic-plots, fig.width=9, fig.height=4.4}
if (!is_fit_error(fit_kernel_fast)) {
  old_par <- par(mfrow = c(1, 2))
  plot(fit_kernel_fast, type = "curve", curve = "level",
       interval = "pointwise", Xdistr = "histogram",
       support.panel = "embedded",
       main = "Kernel level curve (mortality)")
  plot(fit_kernel_fast, type = "curve", curve = "derivative",
       interval = "pointwise", Xdistr = "histogram",
       support.panel = "embedded",
       main = "Kernel derivative curve (mortality)")
  par(old_par)
}
```

Support displays are descriptive. They show where observed values of `G` lie
inside the plotted window; they do not change the stored estimates. This
estimator chapter uses one consistent histogram display, but its titles name
the estimand rather than the display option. Chapter
@sec-visualization shows the alternative density, rug, none, separate-panel,
marker, and contrast plot options.

## Bootstrap Pointwise And Uniform Bands

For kernel inference, `vartype = "robust"` gives analytical pointwise normal
intervals. With `vartype = "bootstrap"`, the package stores bootstrap replicate
curves, pointwise bootstrap percentile intervals, curve covariance matrices,
and finite-grid bootstrap quantile-envelope bands.

The rendered example below keeps the same mortality dataset and the same
zero-anchored reporting grid. In an applied analysis, use many more bootstrap
replications. The band shown by `interval = "both"` is a bootstrap
quantile-envelope band; it is not a studentized sup-$t$ band.

```{r kernel-bootstrap-fit}
kernel_boot_draws <- if (identical(tolower(Sys.getenv("FDID_TUTORIAL_HEAVY")),
                                   "true")) 199L else 19L

fit_kernel_boot <- fit_fdid(
  s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "kernel",
  eval_g = eval_g_kernel,
  h0 = 0.18,
  K_folds = 3,
  vartype = "bootstrap",
  boot = kernel_boot_draws
)

if (!is_fit_error(fit_kernel_boot)) {
  data.frame(
    band_method = fit_kernel_boot$curve_event$simultaneous_band_method,
    effective_bootstrap_draws = fit_kernel_boot$curve_event$delta_uniform_n_boot_eff
  )
}
```

```{r kernel-bootstrap-plots, fig.width=9, fig.height=4.4}
if (!is_fit_error(fit_kernel_boot)) {
  old_par <- par(mfrow = c(1, 2))
  plot(fit_kernel_boot, type = "curve", curve = "level",
       interval = "both", show.uniform.CI = TRUE,
       Xdistr = "histogram", support.panel = "embedded",
       main = "Kernel level bands (mortality)")

  plot(fit_kernel_boot, type = "curve", curve = "derivative",
       interval = "both", show.uniform.CI = TRUE,
       Xdistr = "histogram", support.panel = "embedded",
       main = "Kernel derivative bands (mortality)")
  par(old_par)
}
```

To hide the uniform band while keeping pointwise intervals, use
`show.uniform.CI = FALSE`, mirroring the interflex plotting workflow.

```{r kernel-hide-uniform, fig.width=7, fig.height=4.5}
if (!is_fit_error(fit_kernel_boot)) {
  plot(fit_kernel_boot, type = "curve", curve = "level",
       show.uniform.CI = FALSE,
       Xdistr = "histogram", support.panel = "embedded",
       main = "Kernel pointwise intervals (mortality)")
}
```

## Reporting Contrasts And Derivatives

The curve is useful for diagnosis, but a write-up often needs a smaller number
of quantities. `fdid_contrast()` reports level differences such as
`mu(g1) - mu(g0)`. `fdid_derivative()` reports a derivative at a fixed value of
`G`.

```{r kernel-reporting}
if (!is_fit_error(fit_kernel_boot)) {
  kernel_report <- rbind(
    fdid_contrast(fit_kernel_boot, g0 = min(eval_g_kernel),
                  g1 = max(eval_g_kernel)),
    fdid_derivative(fit_kernel_boot, g0 = stats::median(eval_g_kernel))
  )

  knitr::kable(
    kernel_report[, c("target", "method", "g0", "g1", "estimate",
                      "std.error", "conf.low", "conf.high", "inference")],
    digits = 3
  )
}
```

The fixed-reference coefficient plot gives the same reporting target in a
visual form: choose a reference value of `G`, then display several
`mu(g) - mu(g_ref)` contrasts and their 95% intervals. The reference row is the
algebraic zero anchor.

```{r kernel-contrast-plot, fig.width=7, fig.height=4.5}
if (!is_fit_error(fit_kernel_boot)) {
  plot(fit_kernel_boot, type = "contrast",
       ref.g = 0,
       target.g = eval_g_kernel,
       main = "Kernel contrasts vs G = 0")
}
```

When bootstrap covariance matrices are available, exact-grid contrasts use the
covariance identity `V22 + V11 - 2 * V12`. Otherwise the helper falls back to
replicate curves, stored bands, or pointwise endpoint intervals and labels the
inference source.

## Option Checklist

```{r kernel-option-checklist}
kernel_options <- data.frame(
  option = c("eval_g", "h0", "K_folds", "vartype", "boot",
             "cluster_label", "curve", "interval", "show.uniform.CI",
             "Xdistr", "support.panel", "diff.values"),
  role = c(
    "Evaluation grid; this is part of estimation.",
    "Baseline bandwidth; NULL selects by cross-validation.",
    "Number of folds for bandwidth cross-validation.",
    "Use robust pointwise intervals or bootstrap curve inference.",
    "Bootstrap replications for kernel replicate curves and bands.",
    "Passed to fdid_prepare(); enables cluster bootstrap resampling.",
    "Choose level, derivative, or both curves.",
    "Choose pointwise intervals, uniform bands, both, or none.",
    "Suppress stored uniform bands at plotting time.",
    "Show histogram, density, rug, or no support display.",
    "Draw support inside the curve panel or below it.",
    "Mark reporting values of G on the plot."
  )
)

knitr::kable(kernel_options)
```

For a substantive kernel analysis, the practical checklist is:

- inspect the support of `G` before choosing the grid;
- avoid interpreting sparse tails or large mass points as smooth regions;
- report whether intervals are analytical pointwise or bootstrap pointwise;
- call the simultaneous band a bootstrap quantile-envelope band, not a
  studentized sup-$t$ band;
- report key values with `fdid_contrast()` or `fdid_derivative()` in addition
  to the plot.
