# Kernel Estimator for Continuous G {#sec-kernel-continuous}

```{r setup-kernel, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  message = FALSE,
  warning = FALSE
)
```

The kernel estimator is the transparent first curve workflow for continuous
`G`. It estimates a level curve `mu(g)` and a derivative curve `delta(g)` over
the supplied `eval_g` grid.

## When Kernel Is The Right First Curve

Use `method = "kernel"` when the main object is a smooth curve and you want a
direct view of how the FDID response changes with `G`. It is especially useful
as a first curve because the evaluation grid, bandwidth, and support choices
are explicit.

## Evaluation Grid And Bandwidth

```{r kernel-grid-fit, eval=FALSE}
fit_kernel <- fdid(
  s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "kernel",
  eval_g = eval_g,
  h0 = 0.18,
  vartype = "robust"
)
```

`eval_g` is the grid where level and derivative quantities are estimated.
Leaving `h0 = NULL` requests bandwidth selection; setting `h0` fixes the
bandwidth for reproducibility or sensitivity checks.

## Level Curve

```{r kernel-level, eval=FALSE}
plot(fit_kernel, type = "curve", curve = "level",
     interval = "pointwise", Xdistr = "rug")
```

The level curve is the fitted event-period response `mu(g)`.

## Derivative Curve

```{r kernel-derivative, eval=FALSE}
plot(fit_kernel, type = "curve", curve = "derivative",
     interval = "pointwise", Xdistr = "rug")
```

The derivative curve is the local slope `delta(g)` on the evaluation grid.
Derivative interpretation is more sensitive to bandwidth and support than
level interpretation.

## Pointwise Analytical Intervals

With `vartype = "robust"`, kernel pointwise intervals use stored analytical
standard errors for each grid point. They are pointwise intervals, not uniform
bands.

## Bootstrap Pointwise Intervals

With `vartype = "bootstrap"`, kernel pointwise intervals use stored bootstrap
curve replicates.

```{r kernel-bootstrap, eval=FALSE}
fit_kernel_boot <- fdid(
  s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "kernel",
  eval_g = eval_g,
  h0 = 0.18,
  vartype = "bootstrap",
  boot = 199,
  parallel = FALSE
)
```

Use substantially more replications in a final empirical report than in a fast
tutorial render.

## Bootstrap Quantile-Envelope Bands

Kernel bootstrap bands are finite-grid quantile-envelope bands over stored
bootstrap curves. They should not be described as studentized sup-`t` bands.

```{r kernel-bands, eval=FALSE}
plot(fit_kernel_boot, type = "curve", curve = "level",
     interval = "both", show.uniform.CI = TRUE)
```

## Contrast And Derivative Reporting

Use reporting helpers when the table needs a value, a level contrast, or a
fixed-`G` derivative instead of the whole plot.

```{r kernel-reporting, eval=FALSE}
fdid_contrast(fit_kernel_boot, g0 = eval_g[2], g1 = eval_g[7])
fdid_derivative(fit_kernel_boot, g0 = eval_g[5])
plot(fit_kernel_boot, type = "contrast", ref.g = "median",
     target.type = "quantile", target.g = 5)
```

Exact-grid reporting can use stored covariance or replicate information when
available. Interpolated reporting should be labeled as interpolated.

## Kernel Option Checklist

- Choose an interior `eval_g` grid with enough support.
- Compare at least one bandwidth sensitivity run.
- Label analytical intervals as pointwise.
- Label bootstrap bands as quantile-envelope bands.
- Report fixed-reference contrasts through `fdid_contrast()` or
  `plot(type = "contrast")`.
