# Kernel Estimator for Continuous G {#sec-kernel-continuous}

```{r kernel-setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 7,
  fig.height = 4.8
)
set.seed(20260531)

fit_fdid <- function(...) {
  tryCatch(
    fdid(...),
    error = function(e) {
      structure(list(message = conditionMessage(e)), class = "fdid_fit_error")
    }
  )
}

is_fit_error <- function(x) inherits(x, "fdid_fit_error")

event_table <- function(obj, label) {
  if (is_fit_error(obj)) {
    return(data.frame(
      fit = label, Estimate = NA_real_, Std.Error = NA_real_,
      CI_Lower = NA_real_, CI_Upper = NA_real_, note = obj$message,
      check.names = FALSE
    ))
  }
  out <- obj$est$event
  data.frame(
    fit = label,
    Estimate = as.numeric(out$Estimate[1]),
    Std.Error = as.numeric(out$Std.Error[1]),
    CI_Lower = as.numeric(out$CI_Lower[1]),
    CI_Upper = as.numeric(out$CI_Upper[1]),
    note = "",
    check.names = FALSE
  )
}

kernel_curve_table <- function(obj, n = 8) {
  if (is_fit_error(obj) || is.null(obj$curve_event)) return(data.frame())
  ce <- obj$curve_event
  head(data.frame(
    G = obj$eval_g,
    mu_hat = ce$mu_hat,
    se_mu = ce$se_mu,
    delta_hat = ce$delta_hat,
    se_delta = ce$se_delta
  ), n)
}
```

This chapter gives a focused workflow for `method = "kernel"` with continuous
baseline factor `G`. The structure follows the practical style used in the
`interflex` tutorial and in @liu2025practical: diagnose support first, estimate
the curve, inspect uncertainty, then report specific contrasts or derivatives.

The kernel estimator is useful when the main empirical object is a smooth
continuous-$G$ curve rather than a single slope. It estimates two event-period
objects:

- the level curve, $\mu(g)$, the local FDID response at a value of `G`;
- the derivative curve, $\delta(g)$, the local slope of that response.

Unlike the DML estimators in Chapter @sec-dml-continuous, the kernel estimator
is a fused local smoother: the local linear fit, support weighting, and curve
aggregation are part of one estimator.

## Data Setup And Support

```{r kernel-library-data}
library(fdid)
data(fdid)

mortality$uniqueid <- paste(
  as.character(mortality$provid),
  as.character(mortality$countyid),
  sep = "-"
)

covar <- c("avggrain", "nograin", "urban", "dis_bj", "dis_pc",
           "rice", "minority", "edu", "lnpop")
tr_period <- 1958:1961
ref_period <- 1957

s_cont <- fdid_prepare(
  data = transform(mortality, G_cont = lnpczupu),
  Y_label = "mortality",
  X_labels = covar,
  G_label = "G_cont",
  unit_label = "uniqueid",
  time_label = "year"
)
```

The famine mortality example has a mass point at zero in `lnpczupu`. This is
important because `eval_g` is an estimation grid, not a plotting-only option.
When `eval_g = NULL`, `fdid()` constructs the grid internally:

```r
seq(quantile(G, trim), quantile(G, 1 - trim), length.out = 50)
```

The package default is `trim = 0.05`, so `eval_g = NULL` usually means 50
points from the 5th to the 95th percentile of the observed `G` values after
the estimation sample is formed. `trim` does not delete observations or
winsorize `G`; it only chooses the endpoints of the evaluation grid. If more
than 5 percent of units have `G = 0`, then the lower endpoint is still zero.

In this chapter we deliberately set `trim = 0` while leaving `eval_g = NULL`.
That keeps the convenient 50-point automatic grid but expands it to the full
observed support of `G`. This is useful in the mortality example because the
95%-100% upper tail contains substantively large observed baseline-factor
values rather than a single isolated outlier. Supplying `eval_g` manually
would override this rule and estimate exactly the points supplied.

```{r kernel-support-grid, echo=FALSE}
kernel_package_trim <- 0.05
kernel_chapter_trim <- 0

kernel_package_grid <- seq(
  as.numeric(stats::quantile(s_cont$G, kernel_package_trim, na.rm = TRUE)),
  as.numeric(stats::quantile(s_cont$G, 1 - kernel_package_trim, na.rm = TRUE)),
  length.out = 50
)

kernel_chapter_grid <- seq(
  as.numeric(stats::quantile(s_cont$G, kernel_chapter_trim, na.rm = TRUE)),
  as.numeric(stats::quantile(s_cont$G, 1 - kernel_chapter_trim, na.rm = TRUE)),
  length.out = 50
)

g_positive <- s_cont$G[s_cont$G > 0]
g_tail_95 <- as.numeric(stats::quantile(s_cont$G, 0.95, na.rm = TRUE))

kernel_landmarks <- as.numeric(stats::quantile(
  g_positive,
  probs = c(0.20, 0.50, 0.80),
  na.rm = TRUE
))

support_summary <- data.frame(
  quantity = c("n units", "share G = 0", "package default trim",
               "chapter trim", "package default grid maximum",
               "chapter grid minimum", "chapter grid maximum",
               "chapter grid size", "units above package 95% endpoint",
               "share above package 95% endpoint",
               "20% positive support", "80% positive support"),
  value = c(nrow(s_cont), mean(s_cont$G == 0),
            kernel_package_trim, kernel_chapter_trim,
            max(kernel_package_grid), min(kernel_chapter_grid),
            max(kernel_chapter_grid), length(kernel_chapter_grid),
            sum(s_cont$G > g_tail_95), mean(s_cont$G > g_tail_95),
            min(kernel_landmarks),
            max(kernel_landmarks))
)

knitr::kable(support_summary, digits = 4)
```

The support plot below does not draw all 50 evaluation points. The table records
the grid size, while the figure marks only the full-support upper endpoint, the
package-default 95% endpoint, and a few positive-support landmarks.

```{r kernel-support-plot, fig.width=7, fig.height=5.6}
old_par_support <- par(no.readonly = TRUE)
layout(matrix(c(1L, 2L), ncol = 1), heights = c(4, 1.05))
par(mar = c(4.7, 4.2, 4, 2) + 0.1)
hist(s_cont$G, breaks = 35, col = "gray85", border = "white",
     xlim = c(0, max(kernel_chapter_grid) * 1.02),
     xlab = "Baseline factor G: lnpczupu",
     main = "Observed support for continuous G")
abline(v = max(kernel_package_grid), col = "gray40", lty = 2, lwd = 1.5)
abline(v = kernel_landmarks, col = "firebrick", lty = 3, lwd = 1.5)
points(max(kernel_chapter_grid), 0, pch = 17, col = "steelblue", cex = 1.4)
rug(s_cont$G, col = grDevices::adjustcolor("black", 0.25))
par(mar = c(0, 4.2, 0.2, 2) + 0.1)
plot.new()
legend("center",
       legend = c("full-support upper endpoint", "package trim = 0.05 upper end",
                  "positive-support landmarks"),
       col = c("steelblue", "gray40", "firebrick"), lty = c(NA, 2, 3),
       pch = c(17, NA, NA), lwd = 1.5,
       bty = "o", bg = "white", box.col = "gray75")
layout(1)
par(old_par_support)
```

## Basic Kernel Fit

The next fit uses a fixed bandwidth `h0` so that the tutorial renders quickly
and reproducibly. Leaving `h0 = NULL` asks the package to select the baseline
bandwidth by cross-validation. We leave `eval_g = NULL` but set `trim = 0`, so
the fitted object stores a 50-point grid over the full observed support of `G`.

```{r kernel-basic-fit}
fit_kernel_fast <- fit_fdid(
  s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "kernel",
  h0 = 0.18,
  K_folds = 3,
  trim = kernel_chapter_trim,
  vartype = "robust"
)
```

The first reporting layer is the standard `summary()` method. It gives the
scalar aggregate/event/dynamic rows and then identifies the stored
continuous-$G$ curve object.

```{r kernel-basic-summary}
if (!is_fit_error(fit_kernel_fast)) {
  summary(fit_kernel_fast)
}
```

```{r kernel-basic-grid, echo=FALSE}
if (!is_fit_error(fit_kernel_fast)) {
  data.frame(
    stored_eval_g_points = length(fit_kernel_fast$eval_g),
    stored_grid_min = min(fit_kernel_fast$eval_g),
    stored_grid_max = max(fit_kernel_fast$eval_g)
  )
}
```

The scalar event row is a compact interval-average level-slope summary over the
fitted `eval_g` range. For the kernel estimator, the main objects are still the
stored level and derivative curve arrays. Under `vartype = "robust"`, the
current implementation also stores analytical cross-grid covariance matrices
for those arrays; this is what lets the scalar summary and exact-grid contrasts
use covariance-aware standard errors rather than independent-endpoint
approximations.

```{r kernel-curve-table, echo=FALSE}
if (!is_fit_error(fit_kernel_fast)) {
  knitr::kable(kernel_curve_table(fit_kernel_fast), digits = 3)
}
```

## Level And Derivative Curves

`plot(type = "curve")` draws the stored curve values. Use `curve = "level"` for
the FDID response curve and `curve = "derivative"` for the local slope.

```{r kernel-basic-plots, fig.width=9, fig.height=5.2}
if (!is_fit_error(fit_kernel_fast)) {
  old_par <- par(mfrow = c(1, 2))
  plot(fit_kernel_fast, type = "curve", curve = "level",
       interval = "pointwise", Xdistr = "histogram",
       support.panel = "embedded",
       main = "Kernel level curve (mortality)")
  plot(fit_kernel_fast, type = "curve", curve = "derivative",
       interval = "pointwise", Xdistr = "histogram",
       support.panel = "embedded",
       main = "Kernel derivative curve (mortality)")
  par(old_par)
}
```

Support displays are descriptive. They show where observed values of `G` lie
inside the plotted window; they do not change the stored estimates. This
estimator chapter uses one consistent histogram display, but its titles name
the estimand rather than the display option. Chapter
@sec-visualization shows the alternative density, rug, none, separate-panel,
marker, and contrast plot options.

## Bootstrap Pointwise And Uniform Bands

For kernel inference, keep two routes separate:

| `vartype` | What is stored | Main use |
|---|---|---|
| `"robust"` | analytical pointwise normal intervals plus stacked-sandwich `mu_vcov` and `delta_vcov` over the stored grid | stable pointwise intervals, covariance-aware summaries, contrasts, and derivative values |
| `"bootstrap"` | bootstrap replicate curves, pointwise percentile intervals, bootstrap curve covariance, and finite-grid quantile-envelope bands | pointwise bootstrap intervals and uniform curve displays |

The robust covariance is a cross-grid covariance for the local-linear curve
estimates. Exact-grid level contrasts use the identity
`V22 + V11 - 2 * V12`; off-grid values use the same matrix after linear
interpolation weights. Bootstrap covariance matrices are computed from the
stored replicate curves.

The rendered example below keeps the same mortality dataset and the same
full-support 50-point grid. It uses 200 bootstrap replications so the
pointwise percentile intervals and finite-grid quantile-envelope band are much
less dominated by a handful of extreme resamples than in a quick draft render.
The band shown by `interval = "both"` is a bootstrap quantile-envelope band;
it is not a studentized sup-$t$ band. Even with 200 replications, the
full-support right tail and derivative curve can remain wide because support
is thinner and local slopes are more sensitive than levels.

The metadata row reports the number of bootstrap curves that are usable for the
finite-grid uniform envelope. This count can be smaller than `boot` because the
uniform envelope needs a whole bootstrap curve to be finite at every stored
`eval_g` point. Pointwise percentile intervals are computed column by column
with finite bootstrap draws, but a joint finite-grid envelope drops resampled
curves that fail local-linear estimation at one or more grid points. This is not
an additional trimming or winsorization step; it is a finite-value filter for the
uniform-band calculation.

```{r kernel-bootstrap-fit}
kernel_boot_draws <- 200L

fit_kernel_boot <- fit_fdid(
  s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "kernel",
  h0 = 0.18,
  K_folds = 3,
  trim = kernel_chapter_trim,
  vartype = "bootstrap",
  boot = kernel_boot_draws
)
```

```{r kernel-bootstrap-summary, echo=FALSE}
if (!is_fit_error(fit_kernel_boot)) {
  data.frame(
    band_method = fit_kernel_boot$curve_event$simultaneous_band_method,
    effective_bootstrap_draws = fit_kernel_boot$curve_event$delta_uniform_n_boot_eff
  )
}
```

```{r kernel-bootstrap-plots, fig.width=9, fig.height=5.2}
if (!is_fit_error(fit_kernel_boot)) {
  old_par <- par(mfrow = c(1, 2))
  plot(fit_kernel_boot, type = "curve", curve = "level",
       interval = "both", show.uniform.CI = TRUE,
       Xdistr = "histogram", support.panel = "embedded",
       main = "Kernel level bands (mortality)")

  plot(fit_kernel_boot, type = "curve", curve = "derivative",
       interval = "both", show.uniform.CI = TRUE,
       Xdistr = "histogram", support.panel = "embedded",
       main = "Kernel derivative bands (mortality)")
  par(old_par)
}
```

To hide the uniform band while keeping pointwise intervals, use
`show.uniform.CI = FALSE`, mirroring the interflex plotting workflow.

```{r kernel-hide-uniform, fig.width=7, fig.height=4.5}
if (!is_fit_error(fit_kernel_boot)) {
  plot(fit_kernel_boot, type = "curve", curve = "level",
       show.uniform.CI = FALSE,
       Xdistr = "histogram", support.panel = "embedded",
       main = "Kernel pointwise intervals (mortality)")
}
```

## Reporting Contrasts And Derivatives

The curve is useful for diagnosis, but a write-up often needs a smaller number
of quantities. `fdid_contrast()` reports level differences such as
`mu(g1) - mu(g0)`. `fdid_derivative()` reports a derivative at a fixed value of
`G`.

For these compact reporting examples, we use the analytical `vartype = "robust"`
fit from the basic kernel section rather than the bootstrap object used for the
uniform-band display above. The reporting helper's default `inference = "auto"`
first looks for stored curve covariance; for exact-grid contrasts this is the
`V22 + V11 - 2 * V12` calculation. In this mortality example, however, the
substantively useful reference is the lowest grid value `G = 0`, and the
full-support boundary covariance can be numerically unavailable. We therefore
request `inference = "pointwise"` in the compact table and coefficient plot, so
the output is explicitly labeled as an analytical endpoint approximation.

```{r kernel-reporting, echo=FALSE}
if (!is_fit_error(fit_kernel_fast)) {
  kernel_eval_grid <- fit_kernel_fast$eval_g
  kernel_report <- rbind(
    fdid_contrast(fit_kernel_fast, g0 = min(kernel_eval_grid),
                  g1 = max(kernel_eval_grid), inference = "pointwise"),
    fdid_derivative(fit_kernel_fast, g0 = stats::median(kernel_eval_grid),
                    inference = "pointwise")
  )

  knitr::kable(
    kernel_report[, c("target", "method", "g0", "g1", "estimate",
                      "std.error", "conf.low", "conf.high", "inference")],
    digits = 3
  )
}
```

The fixed-reference coefficient plot gives the same reporting target in a
visual form: choose a reference value of `G`, then display several
`mu(g) - mu(g_ref)` contrasts and their 95% intervals. The reference row is the
algebraic zero anchor.

```{r kernel-contrast-plot, fig.width=7, fig.height=4.5}
if (!is_fit_error(fit_kernel_fast)) {
  plot(fit_kernel_fast, type = "contrast",
       ref.g = "min",
       target.g = "5",
       inference = "pointwise",
       main = "Kernel analytical contrasts vs lowest grid point")
}
```

When the helper is applied to a bootstrap kernel object, exact-grid contrasts
can use the stored bootstrap curve covariance matrix and the covariance identity
`V22 + V11 - 2 * V12`. The robust object shown here now stores the analytical
stacked-sandwich version of the same cross-grid covariance idea when the
relevant grid covariance entries are finite. Use the bootstrap object when the
target is a bootstrap curve display or replicate-based sensitivity check; use
the robust object when the goal is a stable analytical reporting summary, and
read the `inference` column to see whether the helper used covariance or a
labeled pointwise approximation.

## Option Checklist

```{r kernel-option-checklist, echo=FALSE}
kernel_options <- data.frame(
  option = c("eval_g", "h0", "K_folds", "vartype", "boot",
             "cluster_label", "curve", "interval", "show.uniform.CI",
             "Xdistr", "support.panel", "diff.values"),
  role = c(
    "Evaluation grid; NULL uses a 50-point grid controlled by trim.",
    "Baseline bandwidth; NULL selects by cross-validation.",
    "Number of folds for bandwidth cross-validation.",
    "Use robust pointwise intervals or bootstrap curve inference.",
    "Bootstrap replications for kernel replicate curves and bands.",
    "Passed to fdid_prepare(); enables cluster bootstrap resampling.",
    "Choose level, derivative, or both curves.",
    "Choose pointwise intervals, uniform bands, both, or none.",
    "Suppress stored uniform bands at plotting time.",
    "Show histogram, density, rug, or no support display.",
    "Draw support inside the curve panel or below it.",
    "Mark reporting values of G on the plot."
  )
)

knitr::kable(kernel_options)
```

For a substantive kernel analysis, the practical checklist is:

- inspect the support of `G` before choosing the grid;
- avoid interpreting sparse tails or large mass points as smooth regions;
- report whether intervals are analytical pointwise, bootstrap pointwise, or
  covariance-based contrasts from stored curve covariance;
- call the simultaneous band a bootstrap quantile-envelope band, not a
  studentized sup-$t$ band;
- report key values with `fdid_contrast()` or `fdid_derivative()` in addition
  to the plot.
