# DML Estimators for Continuous G {#sec-dml-continuous}

```{r dml-setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 7,
  fig.height = 4.8
)
set.seed(20260601)

fit_fdid <- function(...) {
  tryCatch(
    fdid(...),
    error = function(e) {
      structure(list(message = conditionMessage(e)), class = "fdid_fit_error")
    }
  )
}

is_fit_error <- function(x) inherits(x, "fdid_fit_error")

event_row <- function(obj, label, target, learner = NA_character_,
                      density_method = NA_character_,
                      signal_map = NA_character_,
                      truth = NA_real_) {
  if (is_fit_error(obj)) {
    return(data.frame(
      fit = label, target = target, learner = learner,
      density_method = density_method, signal_map = signal_map,
      truth = truth, Estimate = NA_real_, Std.Error = NA_real_,
      CI_Lower = NA_real_, CI_Upper = NA_real_,
      note = obj$message,
      check.names = FALSE
    ))
  }
  ev <- obj$est$event
  data.frame(
    fit = label, target = target, learner = learner,
    density_method = density_method, signal_map = signal_map,
    truth = truth,
    Estimate = as.numeric(ev$Estimate[1]),
    Std.Error = as.numeric(ev$Std.Error[1]),
    CI_Lower = as.numeric(ev$CI_Lower[1]),
    CI_Upper = as.numeric(ev$CI_Upper[1]),
    note = "",
    check.names = FALSE
  )
}

metadata_table <- function(obj) {
  if (is_fit_error(obj) || is.null(obj$dml_metadata)) return(data.frame())
  vals <- vapply(obj$dml_metadata, function(z) {
    if (length(z) == 0L || is.null(z)) return(NA_character_)
    paste(as.character(z), collapse = ", ")
  }, character(1L))
  data.frame(field = names(vals), value = unname(vals), row.names = NULL)
}
```

This chapter gives a focused workflow for the double/debiased machine learning
(DML) estimators in `fdid` when `G` is continuous. To keep the target logic
clear, the entire chapter uses one synthetic panel with a known treatment
effect curve. The same data-generating process carries the scalar slope,
flexible curve, curve-inference, reporting, and incremental-derivative
examples.

The most important rule is target-first: choose the estimand before choosing
the method.

## Target Map

```{r dml-target-map, echo=FALSE}
dml_targets <- data.frame(
  method = c("dml_plr", "dml_flex", "dml_incremental"),
  target = c(
    "Partially linear scalar slope",
    "Flexible level curve theta(g) and derivative theta'(g)",
    "Observed-population average derivative E[partial_g mu(G, X)]"
  ),
  when_to_use = c(
    "You want one adjusted slope summary.",
    "You want a level curve, derivative curve, or fixed-G contrast.",
    "You want the average local derivative over the observed G distribution."
  ),
  main_output = c(
    "event summary row",
    "curve object plus plot/reporting helpers",
    "event summary row"
  )
)

knitr::kable(dml_targets)
```

All three methods work with the FDID first-difference outcome

$$
\Delta Y_i = \bar Y_{i,\mathrm{event}} - Y_{i,\mathrm{ref}}.
$$

DML then proceeds in three steps:

1. Learn nuisance functions with cross-fitting.
2. Construct an orthogonal signal for the target.
3. Aggregate or map that signal into the requested object: a scalar slope, a
   smooth curve in `G`, or an observed-population average derivative.

Binary `method = "dml_binary"` is covered in Chapter @sec-binary because it
targets a binary contrast, not a continuous-`G` scalar or curve.

## Synthetic Data Setup

The DGP has four periods. Period 2 is the reference period and periods 3 and 4
are the event window. Each unit has a continuous baseline factor
$G_i \sim \mathrm{Beta}(4, 4)$ and three covariates. This gives dense
interior support and avoids letting a few extreme endpoints dominate a
tutorial figure. The event-period effect is

$$
\tau(g) = 0.4 + 0.6g + 0.8g^2.
$$

Period 3 receives $1\times \tau(G_i)$ and period 4 receives
$2\times \tau(G_i)$. Therefore the event-average level curve is
$1.5\tau(g)$, the derivative curve is $1.5\tau'(g)$, and the observed
incremental target is the sample average of $1.5\tau'(G_i)$.

```{r dml-synthetic-data}
library(fdid)

n_dml_sim <- 300
time_dml_sim <- 1:4
G_dml_sim <- rbeta(n_dml_sim, shape1 = 4, shape2 = 4)
X1_dml_sim <- rnorm(n_dml_sim)
X2_dml_sim <- rnorm(n_dml_sim)
X3_dml_sim <- rbinom(n_dml_sim, size = 1, prob = 0.45)

dml_tau <- function(g) 0.4 + 0.6 * g + 0.8 * g^2
dml_tau_deriv <- function(g) 0.6 + 1.6 * g
dml_true_level <- function(g) 1.5 * dml_tau(g)
dml_true_deriv <- function(g) 1.5 * dml_tau_deriv(g)

dml_sim_panel <- data.frame(
  id = rep(seq_len(n_dml_sim), each = length(time_dml_sim)),
  time = rep(time_dml_sim, times = n_dml_sim),
  G = rep(G_dml_sim, each = length(time_dml_sim)),
  x1 = rep(X1_dml_sim, each = length(time_dml_sim)),
  x2 = rep(X2_dml_sim, each = length(time_dml_sim)),
  x3 = rep(X3_dml_sim, each = length(time_dml_sim))
)

dml_sim_panel$Y <- with(
  dml_sim_panel,
  0.35 * x1 - 0.25 * x2 + 0.20 * x3 +
    0.10 * sin(2 * pi * G) + 0.05 * time +
    ifelse(time >= 3, (time - 2) * dml_tau(G), 0) +
    rnorm(nrow(dml_sim_panel), sd = 0.08)
)

s_dml_sim <- fdid_prepare(
  data = dml_sim_panel,
  Y_label = "Y",
  X_labels = c("x1", "x2", "x3"),
  G_label = "G",
  unit_label = "id",
  time_label = "time"
)
```

```{r dml-dgp-summary, echo=FALSE}
dml_true_incremental <- mean(dml_true_deriv(G_dml_sim))
dml_default_grid <- seq(
  as.numeric(stats::quantile(s_dml_sim$G, 0.05, na.rm = TRUE)),
  as.numeric(stats::quantile(s_dml_sim$G, 0.95, na.rm = TRUE)),
  length.out = 50
)

dgp_summary <- data.frame(
  quantity = c("units", "periods", "reference period", "event periods",
               "default curve grid size", "default grid minimum",
               "default grid maximum", "true observed average derivative"),
  value = c(n_dml_sim, length(time_dml_sim), 2, "3, 4",
            length(dml_default_grid), min(dml_default_grid),
            max(dml_default_grid), dml_true_incremental)
)

knitr::kable(dgp_summary, digits = 3)
```

## Common DML Inputs

The rendered examples use `learner = "grf"` to make nuisance learning flexible.
The package default remains `learner = "linear"` because it is faster and has
fewer backend requirements. GRF examples take longer because the nuisance
models are fit inside cross-fitting loops.

```{r dml-common-inputs, echo=FALSE}
dml_inputs <- data.frame(
  option = c("learner", "K", "S", "eval_g", "trim",
             "density_method", "signal_map", "dml_inference"),
  role = c(
    "Nuisance learner; this chapter uses grf.",
    "Number of cross-fitting folds.",
    "Number of sample splits.",
    "Curve evaluation grid; only used by dml_flex.",
    "Default eval_g endpoint rule for dml_flex.",
    "Conditional-density approximation for dml_flex and dml_incremental.",
    "Second-stage curve mapping for dml_flex only.",
    "Scalar DML inference style for dml_plr."
  )
)

knitr::kable(dml_inputs)
```

## Scalar PLR Benchmark

`method = "dml_plr"` estimates one partially linear slope. It is a benchmark
for the question "is there one average adjusted relationship between
continuous `G` and the FDID first-difference outcome?"

```{r dml-plr-fit}
fit_plr_score <- fit_fdid(
  s_dml_sim,
  tr_period = 3:4,
  ref_period = 2,
  method = "dml_plr",
  learner = "grf",
  K = 3,
  S = 1,
  dml_inference = "score"
)

fit_plr_multiplier <- fit_fdid(
  s_dml_sim,
  tr_period = 3:4,
  ref_period = 2,
  method = "dml_plr",
  learner = "grf",
  K = 3,
  S = 1,
  dml_inference = "score_multiplier",
  dml_boot = 99
)
```

```{r dml-plr-table, echo=FALSE}
plr_table <- rbind(
  event_row(fit_plr_score, "PLR score", "partially linear slope",
            learner = "grf"),
  event_row(fit_plr_multiplier, "PLR score multiplier",
            "partially linear slope", learner = "grf")
)

knitr::kable(plr_table[, c("fit", "target", "learner", "Estimate",
                           "Std.Error", "CI_Lower", "CI_Upper", "note")],
             digits = 3)
```

The default scalar DML interval is analytical score inference. The
`score_multiplier` option changes the critical value using a scalar multiplier
process based on stored score contributions. It is not a nonparametric refit
bootstrap.

```{r dml-plr-metadata, echo=FALSE}
if (!is_fit_error(fit_plr_multiplier)) {
  knitr::kable(metadata_table(fit_plr_multiplier))
}
```

## Flexible DML Curves

`method = "dml_flex"` targets the flexible level curve

$$
\theta(g) = E_X[\mu(g, X)]
$$

and its derivative with respect to `G`. The implementation cross-fits nuisance
models, builds orthogonal pseudo-outcomes, and maps those pseudo-outcomes back
to the `eval_g` grid.

Two option families are easy to confuse:

- `density_method` estimates the conditional density
  $f_{G \mid X}(g \mid x)$ needed by the continuous-treatment score. It is
  used by `dml_flex` and `dml_incremental`.
- `signal_map` maps the cross-fitted `dml_flex` signal back to the `eval_g`
  grid. It is used by `dml_flex` only.

`signal_map = "blp_spline"` is therefore not a density algorithm. It is a
B-spline best linear projection mapping path that stores finite-grid
covariance matrices for level and derivative curves.

```{r dml-density-table, echo=FALSE}
density_table <- data.frame(
  density_method = c("residual_kde", "location_scale", "local_kde"),
  algorithm = c(
    "Model E[G | X], then KDE residuals.",
    "Model E[G | X] and residual scale, then KDE standardized residuals.",
    "Product kernel over covariates plus Gaussian KDE over G."
  ),
  practical_reading = c(
    "Fast default; best for location-shift designs.",
    "Allows heteroskedastic G | X.",
    "Most nonparametric; sensitive to covariate dimension."
  )
)

knitr::kable(density_table)
```

```{r dml-flex-fit}
dml_curve_draws <- 99L

fit_flex_spline <- fit_fdid(
  s_dml_sim,
  tr_period = 3:4,
  ref_period = 2,
  method = "dml_flex",
  learner = "grf",
  K = 3,
  S = 1,
  signal_map = "spline",
  density_method = "residual_kde",
  boot = dml_curve_draws
)

fit_flex_blp <- fit_fdid(
  s_dml_sim,
  tr_period = 3:4,
  ref_period = 2,
  method = "dml_flex",
  learner = "grf",
  K = 3,
  S = 1,
  signal_map = "blp_spline",
  density_method = "location_scale",
  boot = dml_curve_draws
)
```

```{r dml-flex-summary, echo=FALSE}
flex_table <- rbind(
  event_row(fit_flex_spline, "dml_flex spline", "average level summary",
            learner = "grf", density_method = "residual_kde",
            signal_map = "spline"),
  event_row(fit_flex_blp, "dml_flex BLP spline", "average level summary",
            learner = "grf", density_method = "location_scale",
            signal_map = "blp_spline")
)

knitr::kable(flex_table[, c("fit", "signal_map", "density_method",
                            "Estimate", "Std.Error", "CI_Lower", "CI_Upper",
                            "note")],
             digits = 3)

if (!is_fit_error(fit_flex_spline)) {
  data.frame(
    stored_eval_g_points = length(fit_flex_spline$eval_g),
    stored_grid_min = min(fit_flex_spline$eval_g),
    stored_grid_max = max(fit_flex_spline$eval_g)
  )
}
```

The scalar event row for `dml_flex` is a compatibility summary. The curve
fields are the primary output, so the table below prints a compact excerpt
rather than all 50 grid rows.

```{r dml-flex-curve-table, echo=FALSE}
if (!is_fit_error(fit_flex_spline)) {
  flex_curve <- data.frame(
    G = fit_flex_spline$eval_g,
    theta_hat = fit_flex_spline$curve_event$theta_hat,
    true_theta = dml_true_level(fit_flex_spline$eval_g),
    se_theta = fit_flex_spline$curve_event$se_theta,
    delta_hat = fit_flex_spline$curve_event$delta_hat,
    true_delta = dml_true_deriv(fit_flex_spline$eval_g),
    se_delta = fit_flex_spline$curve_event$se_delta
  )
  flex_curve_excerpt <- flex_curve[
    unique(round(seq(1, nrow(flex_curve), length.out = 8))),
    ,
    drop = FALSE
  ]
  knitr::kable(flex_curve_excerpt, digits = 3)
}
```

## DML Curve Inference

The DML curve inference labels are method-specific.

- Pointwise intervals for `dml_flex` are second-stage mapping intervals.
- With `signal_map = "spline"`, the simultaneous band is a practical residual
  multiplier band over cross-fitted signal-mapping residuals.
- With `signal_map = "blp_spline"`, the object stores spline-basis covariance
  matrices and uses a Gaussian max-$t$ critical value for finite-grid bands
  when the covariance is usable.
- Neither path is a full nonparametric DML refit bootstrap over all nuisance
  stages.

```{r dml-flex-inference-metadata, echo=FALSE}
if (!is_fit_error(fit_flex_blp)) {
  inference_rows <- data.frame(
    field = c("signal_map", "density_method", "band_method",
              "theta_vcov_available", "delta_vcov_available",
              "gaussian_crit_theta", "simultaneous_crit_theta"),
    value = c(
      fit_flex_blp$signal_map,
      fit_flex_blp$density_method,
      fit_flex_blp$curve_event$simultaneous_band_method,
      !is.null(fit_flex_blp$curve_event$theta_vcov),
      !is.null(fit_flex_blp$curve_event$delta_vcov),
      round(fit_flex_blp$curve_event$gaussian_crit_theta, 3),
      round(fit_flex_blp$curve_event$simultaneous_crit_theta, 3)
    )
  )
  knitr::kable(inference_rows)
}
```

The firebrick dashed line in the figure below is the known event-average truth
from the DGP, evaluated on the stored 50-point default grid.

```{r dml-synthetic-ci-plots, fig.width=9, fig.height=5.2}
if (!is_fit_error(fit_flex_blp)) {
  dml_grid_blp <- fit_flex_blp$eval_g
  true_level <- dml_true_level(dml_grid_blp)
  true_deriv <- dml_true_deriv(dml_grid_blp)

  old_par <- par(mfrow = c(1, 2))
  plot(fit_flex_blp, type = "curve", curve = "level",
       interval = "both", Xdistr = "histogram",
       support.panel = "embedded",
       main = "Synthetic DML level with CI")
  lines(dml_grid_blp, true_level, col = "firebrick", lty = 2, lwd = 2)

  plot(fit_flex_blp, type = "curve", curve = "derivative",
       interval = "both", Xdistr = "histogram",
       support.panel = "embedded",
       main = "Synthetic DML derivative with CI")
  lines(dml_grid_blp, true_deriv, col = "firebrick", lty = 2, lwd = 2)
  par(old_par)
}
```

## Reporting Curve Quantities

Use the reporting helpers when the estimand is a specific level contrast or
derivative value. For BLP-spline curves, `fdid_contrast()` can use the stored
curve covariance matrix, so exact-grid contrasts reduce to the familiar
`V22 + V11 - 2 * V12` identity.

```{r dml-reporting-calls, eval=FALSE}
fdid_contrast(fit_flex_blp, g0 = min(fit_flex_blp$eval_g),
              g1 = max(fit_flex_blp$eval_g))
fdid_derivative(fit_flex_blp, g0 = stats::median(fit_flex_blp$eval_g))
```

```{r dml-reporting, echo=FALSE}
if (!is_fit_error(fit_flex_blp)) {
  dml_eval_grid <- fit_flex_blp$eval_g
  dml_report <- rbind(
    fdid_contrast(fit_flex_blp, g0 = min(dml_eval_grid),
                  g1 = max(dml_eval_grid)),
    fdid_derivative(fit_flex_blp, g0 = stats::median(dml_eval_grid))
  )

  knitr::kable(
    dml_report[, c("target", "method", "g0", "g1", "estimate",
                   "std.error", "conf.low", "conf.high", "inference")],
    digits = 3
  )
}
```

## Incremental Average Derivative

`method = "dml_incremental"` is not the average of the `dml_flex` derivative
curve. It targets the observed-population average derivative directly, using
an orthogonal score that includes the conditional density score for `G | X`.
In this synthetic panel, the true target is the sample average of
$1.5\tau'(G_i)$.

```{r dml-incremental-fit}
fit_inc_residual <- fit_fdid(
  s_dml_sim,
  tr_period = 3:4,
  ref_period = 2,
  method = "dml_incremental",
  learner = "grf",
  K = 3,
  S = 1,
  density_method = "residual_kde"
)

fit_inc_location <- fit_fdid(
  s_dml_sim,
  tr_period = 3:4,
  ref_period = 2,
  method = "dml_incremental",
  learner = "grf",
  K = 3,
  S = 1,
  density_method = "location_scale"
)

fit_inc_local <- fit_fdid(
  s_dml_sim,
  tr_period = 3:4,
  ref_period = 2,
  method = "dml_incremental",
  learner = "grf",
  K = 3,
  S = 1,
  density_method = "local_kde"
)
```

```{r dml-incremental-table, echo=FALSE}
incremental_table <- rbind(
  event_row(fit_inc_residual, "incremental residual KDE",
            "observed average derivative", learner = "grf",
            density_method = "residual_kde", truth = dml_true_incremental),
  event_row(fit_inc_location, "incremental location-scale",
            "observed average derivative", learner = "grf",
            density_method = "location_scale", truth = dml_true_incremental),
  event_row(fit_inc_local, "incremental local KDE",
            "observed average derivative", learner = "grf",
            density_method = "local_kde", truth = dml_true_incremental)
)

knitr::kable(incremental_table[, c("fit", "density_method", "truth",
                                   "Estimate", "Std.Error", "CI_Lower",
                                   "CI_Upper", "note")],
             digits = 3)
```

## Learner Choice

The DML estimators support several nuisance learners. Optional learner
packages are loaded only when requested, so a tutorial should check
availability before running a large grid.

```{r dml-learner-availability, echo=FALSE}
learner_package <- c(
  linear = NA_character_,
  gam = "mgcv",
  lasso = "glmnet",
  ridge = "glmnet",
  elasticnet = "glmnet",
  ranger = "ranger",
  grf = "grf",
  nnet = "nnet",
  xgboost = "xgboost"
)

learner_available <- function(learner) {
  pkg <- learner_package[[learner]]
  is.na(pkg) || requireNamespace(pkg, quietly = TRUE)
}

learner_table <- data.frame(
  learner = names(learner_package),
  package = unname(learner_package),
  available = vapply(names(learner_package), learner_available, logical(1)),
  stringsAsFactors = FALSE
)

knitr::kable(learner_table)
```

## High-Dimensional Expansions

High-dimensional expansions are another sensitivity dimension. The example
below creates a polynomial and interaction basis from the same synthetic
prepared data. This is a design step; for a full analysis, vary one dimension
at a time: learner, expanded design, `density_method`, then `signal_map`.

```{r dml-hd-expand}
s_dml_sim_hd <- fdid_expand_covariates(
  s_dml_sim,
  basis_type = "polynomial",
  poly_degree = 2,
  include_interactions = TRUE
)
```

```{r dml-hd-design, echo=FALSE}
hd_table <- data.frame(
  design = c("raw covariates", "polynomial plus interactions"),
  covariate_columns = c(
    length(grep("^x[0-9]+$", names(s_dml_sim))),
    length(grep("^x[0-9]+$", names(s_dml_sim_hd)))
  )
)

knitr::kable(hd_table)
```

## DML Checklist

Before reporting a DML result, check:

- whether the target is scalar, curve, or incremental;
- whether `G` is truly continuous or binary;
- which nuisance learner was used and whether optional packages were loaded;
- which conditional-density method was used for `dml_flex` or
  `dml_incremental`;
- which curve mapping was used for `dml_flex`;
- whether the interval is analytical score, scalar score multiplier,
  practical signal-residual multiplier, or BLP-spline Gaussian max-$t$;
- whether curve conclusions rely on sparse support or a fragile derivative.
