# DML Estimators for Continuous G {#sec-dml-continuous}

```{r setup-dml, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  message = FALSE,
  warning = FALSE
)
```

DML workflows in `fdid` are target-specific orthogonal-score workflows. Do not
treat them as interchangeable estimators for one quantity.

## DML Targets

| Method | Target | Main output |
|---|---|---|
| `dml_plr` | Scalar partially linear dose slope | `summary()` |
| `dml_flex` | Flexible level curve `theta(g)` and derivative `theta_prime(g)` | curve plots, `fdid_contrast()`, `fdid_derivative()` |
| `dml_incremental` | Observed-population average derivative | `summary()` and density diagnostics |
| `dml_binary` | Binary-`G` AIPW/IRM contrast | `summary()` |

All continuous-treatment DML paths use the FDID first-difference outcome
`Delta Y_i = mean(Y_i, event) - Y_i, ref` conceptually, then apply
cross-fitted nuisance learning and target-specific orthogonal scores.

## `dml_plr`: Scalar Linear Dose Slope

```{r dml-plr, eval=FALSE}
fit_plr <- fdid(
  s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "dml_plr",
  learner = "linear",
  K = 3,
  S = 1,
  dml_inference = "score"
)
summary(fit_plr)
```

`dml_plr` is a scalar benchmark. It should not be described as the whole
continuous-`G` curve.

## `dml_flex`: Flexible Level And Derivative Curves

```{r dml-flex, eval=FALSE}
fit_flex <- fdid(
  s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "dml_flex",
  learner = "linear",
  eval_g = eval_g,
  K = 3,
  S = 1,
  density_method = "residual_kde",
  signal_map = "spline",
  boot = 199
)

plot(fit_flex, type = "curve", curve = "level")
plot(fit_flex, type = "curve", curve = "derivative")
```

`dml_flex` targets a flexible level curve and derivative curve after nuisance
adjustment. Derivative estimates are usually more fragile than level estimates.

## `dml_incremental`: Observed Average Derivative

```{r dml-incremental, eval=FALSE}
fit_incremental <- fdid(
  s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "dml_incremental",
  learner = "linear",
  K = 3,
  S = 1,
  density_method = "residual_kde"
)
summary(fit_incremental)
```

`dml_incremental` reports an observed-population average derivative. It is not
merely the average of a plotted `dml_flex` derivative curve.

## Learners, Cross-Fitting, And Repeated Splits

`learner` controls nuisance models. Linear learners are fast and transparent;
lasso, ridge, elasticnet, ranger, grf, gam, nnet, and xgboost are available
when their packages are installed. `K` controls cross-fitting folds and `S`
controls repeated splits. Use small values for exploratory checks and larger
values for final robustness when runtime allows.

## Conditional Density Choices

`density_method` controls the conditional density `f_{G|X}(g|x)` for
`dml_flex` and `dml_incremental`.

| `density_method` | Practical role |
|---|---|
| `residual_kde` | Fast default; KDE of residualized `G` |
| `location_scale` | Allows heteroskedastic `G | X` through a scale model |
| `local_kde` | More local and more sensitive to covariate dimension |

Treat density choice as a sensitivity dimension.

## BLP Spline Is A Signal Map, Not A Density Estimator

`signal_map` controls how the cross-fitted `dml_flex` signal is mapped to the
evaluation grid. `signal_map = "blp_spline"` is a B-spline best linear
projection signal map. It is not a `density_method`.

## DML Inference Labels

- Scalar DML robust intervals come from orthogonal-score variance.
- Scalar score-multiplier intervals use multiplier draws over stored influence
  values.
- Non-BLP `dml_flex` bands are practical signal-mapping multiplier bands.
- `signal_map = "blp_spline"` stores finite-grid covariance matrices and can
  use Gaussian max-`t` bands when usable.

## DML Diagnostics

Check signal summaries, density diagnostics, low-density shares, derivative
nuisance diagnostics, cluster counts when clustering is used, and repeated
split stability. Diagnostics are part of the DML result object and should be
reported when they explain a substantive choice.
