# Visualization Options {#sec-visualization}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE, message = FALSE, warning = FALSE)
```

In this chapter, we explore various visualization options available in the `fdid` package.

```{r}
library(fdid)
```

```{r readcsv, include=FALSE}
data(fdid)
```

```{r gen, include=FALSE}
mortality$uniqueid <- paste(as.character(mortality$provid), as.character(mortality$countyid), sep = "-")
mortality$highzupu <- ifelse(mortality$pczupu >= median(mortality$pczupu, na.rm = TRUE), 1, 0)
```

```{r parameters, include=FALSE}
tr_period <- 1958:1961
ref_period <- 1957
treat <- "highzupu"
Y <- "mortality"
time <- "year"
unit <- "uniqueid"
covar <- c("avggrain", "nograin", "urban", "dis_bj", "dis_pc",
           "rice", "minority", "edu", "lnpop")
```

```{r prepare, include=FALSE}
prepared_d <- fdid_prepare(
  data = mortality,
  Y_label = Y,
  X_labels = covar,
  G_label = treat,
  unit_label = unit,
  time_label = time
)
```

```{r ols1, include=FALSE}
fdid_results_ols1 <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  vartype = "robust"
)
```

```{r ols2, include=FALSE}
fdid_results_ols2 <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "ols2",
  vartype = "robust"
)
```

```{r did, include=FALSE}
fdid_results_did <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "did",
  vartype = "robust"
)
```

```{r ebal, cache=TRUE, include=FALSE}
fdid_results_ebal <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "ebal",
  vartype = "robust",
  target.pop = "1"
)
```

```{r ipw, cache=TRUE, include=FALSE}
fdid_results_ipw <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "ipw",
  vartype = "robust"
)
```

```{r aipw, cache=TRUE, include=FALSE}
fdid_results_aipw <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "aipw",
  vartype = "robust"
)
```

```{r boot, cache=TRUE, include=FALSE}
fdid_results_ols1_boot <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  vartype = "bootstrap",
  parallel = FALSE,
  nsims = 19
)
```

```{r jackknife, cache=TRUE, include=FALSE}
fdid_results_ols1_jack <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  vartype = "jackknife",
  parallel = FALSE
)
```

```{r ols2_target1, cache=TRUE, include=FALSE}
fdid_results_ols2_target_1 <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "ols2",
  vartype = "robust",
  target.pop = "1"
)
```


## Raw means

-  `ci = FALSE` removes the 95% confidence intervals.

```{r}
plot(fdid_results_ols1, type = "raw", ci = FALSE)
```

-  `connected = TRUE` connects points with lines (default: `FALSE`).

```{r}
plot(fdid_results_ols1, type = "raw", connected = TRUE)
```

-  `group_labels` renames the two groups.

```{r}
plot(fdid_results_ols1, type = "raw", connected = TRUE,
     group_labels = c("Low Social Capital", "High Social Capital"))
```

-   `palette` chooses an RColorBrewer palette (default: "Set2").

```{r}
plot(fdid_results_ols1, type = "raw", connected = TRUE,
     group_labels = c("Low Social Capital", "High Social Capital"), palette = "Set1")
```

-   `shade_periods` controls the position of shaded region. Default is `x$tr_period`, i.e. the event window.
    - By setting `shade_periods = c(1954, 1957)`, the pre-event periods can be set shaded.
    - By setting `shade_periods = NULL`, the user can also remove the shaded region.

```{r}
plot(fdid_results_ols1, type = "raw", connected = TRUE,
     group_labels = c("Low Social Capital", "High Social Capital"),
     shade_periods = c(1954, 1957))
```

```{r}
plot(fdid_results_ols1, type = "raw", connected = TRUE,
     group_labels = c("Low Social Capital", "High Social Capital"),
     shade_periods = NULL)
```

-   `alpha_shade` controls the transparency of the shaded region.

```{r}
plot(fdid_results_ols1, type = "raw", connected = TRUE,
     group_labels = c("Low Social Capital", "High Social Capital"),
     alpha_shade = 0.6)
```

-   The user can also set `ylim`, `xlab`, `ylab`, and `main` (currently, xlim is not supported).

```{r}
plot(fdid_results_ols1, type = "raw", connected = TRUE,
     group_labels = c("Low Social Capital", "High Social Capital"),
     ylim = c(0, 80), xlab = "Year", ylab = "Mortality", main = "Raw Mortality")
```

## Dynamic estimates

-   `ci = FALSE` removes the 95% confidence intervals.

```{r}
plot(fdid_results_ols1, type = "dynamic", ci = FALSE)
```

-   `connected = TRUE` connects points with lines (default: `FALSE`).

```{r}
plot(fdid_results_ols1, type = "dynamic", connected = TRUE)
```

-   `shade_periods`, `alpha_shade`, `ylim`, `xlab`, `ylab`, and `main` customize the plot as above.

```{r}
plot(fdid_results_ols1, type = "dynamic", connected = TRUE,
     shade_periods = c(1954, 1957),
     alpha_shade = 0.8, ylim = c(-10, 4),
     ylab = "Estimates", xlab = "Year",
     main = "Effect Modification of High Social Capital"
     )
```

-   `xlim` zooms into a subset of years.

```{r}
plot(fdid_results_ols1, type = "dynamic",
     alpha_shade = 0.8, ylim = c(-10, 4), xlim = c(1954, 1961),
     ylab = "Estimates", xlab = "Year",
     main = "Hiding Post-Event Periods"
     )
```

## Overlap test

The user can adjust `palette`, `ylim`, `xlab`, and `main`:

```{r}
plot(fdid_results_aipw, type = "overlap",
     palette = "Set1", ylim = c(-0.1, 0.1),
     xlab = 'Score', main = 'Diagnostic Test for Overlap Assumption'
     )
```

## Continuous-G curves

For `method = "kernel"` and `method = "dml_flex"`, `type = "curve"` plots the
event-period curve over the evaluation grid for $G$.

The evaluation grid is part of estimation. If `eval_g` is supplied, `kernel`
and `dml_flex` estimate curve values at exactly those grid points; if it is
left as `NULL`, the package builds its default 50-point grid from
`quantile(G, trim)` to `quantile(G, 1 - trim)`. The curve plot displays the
stored `eval_g` rows and does not estimate additional points while plotting.
The `trim` argument controls this default grid construction only; it is not a
sample-trimming or winsorization rule.

Use the `curve` argument to choose the estimand. The only method-specific
default is `curve = "auto"`: kernel plots default to the derivative curve,
while `dml_flex` plots default to the level curve. Explicitly setting
`curve = "level"` or `curve = "derivative"` makes the visual comparison
transparent.

| Method | Level curve | Derivative curve |
|---|---|---|
| `"kernel"` | `curve = "level"` | `curve = "derivative"` |
| `"dml_flex"` | `curve = "level"` | `curve = "derivative"` |

`method = "dml_incremental"` is scalar; it reports the observed-population
average derivative through `summary()` rather than `type = "curve"`.

For numeric reporting from these stored curves, use `fdid_contrast()` for
level comparisons and `fdid_derivative()` for fixed-$g$ derivative values.

The main continuous-G plot controls are:

| Argument | Options | What it controls |
|---|---|---|
| `curve` | `"auto"`, `"level"`, `"derivative"`, `"both"` | which stored curve is drawn |
| `interval` | `"auto"`, `"pointwise"`, `"uniform"`, `"both"`, `"none"` | pointwise intervals, simultaneous bands, both, or neither |
| `show.uniform.CI` | `TRUE`, `FALSE` | whether stored simultaneous bands are eligible to draw |
| `Xdistr` | `"histogram"`, `"density"`, `"rug"`, `"none"` | observed support display for continuous `G` |
| `support.panel` | `"embedded"`, `"separate"` | whether support is drawn inside the curve panel or below it |
| `show.eval_g` | `TRUE`, `FALSE` | ticks for the stored evaluation grid |
| `legend.position` | `"auto"`, `"topright"`, `"topleft"`, `"bottomright"`, `"bottomleft"`, `"none"` | place or suppress plot legends; `"auto"` draws a boxed legend below the plot |
| `diff.values`, `contrast.values` | numeric vector | vertical reference lines at selected `G` values |
| `line.color`, `line.size`, `ci.color`, `ci.alpha`, `band.color`, `band.lty`, `hist.color`, `density.color`, `support.alpha` | graphics values | base-graphics style controls |
| `xlim`, `ylim`, `xlab`, `ylab`, `main` | graphics values | axis limits and labels |

For fixed-reference contrast coefficient plots, use `type = "contrast"`.
The main additional controls are:

| Argument | Options | What it controls |
|---|---|---|
| `ref.g` | numeric value, `"median"`, `"mean"`, `"min"`, `"max"`, `"first"` | reference level $g_{\mathrm{ref}}$; default is `"median"` for level mode and the first quantile target for quantile mode |
| `include.ref` | `TRUE`, `FALSE` | whether to draw the algebraic zero reference row |
| `target.type` | `"level"`, `"quantile"` | whether `target.g` gives absolute `G` levels or quantile targets |
| `target.g` | numeric vector, integer count, `"q5"`, or `"5"` shorthand | target levels, or quantile request when `target.type = "quantile"` |
| `contrast.values` | numeric vector | alias for level targets when `type = "contrast"`; cannot be combined with `target.g` |
| `n.contrast`, `max.contrast` | positive integers | automatic displayed row count and cap, including the reference row when `include.ref = TRUE` |
| `quantile.source` | `"support"`, `"eval_g"` | whether quantile targets use observed `G` support within the fitted range or the fitted grid itself |
| `inference` | `"auto"`, `"replicate"`, `"vcov"`, `"band"`, `"pointwise"` | rule passed to `fdid_contrast()` for each displayed coefficient |
| `interpolate` | `"linear"`, `"none"` | whether off-grid `ref.g` or target values are linearly interpolated |
| `ci` | `TRUE`, `FALSE` | whether coefficient CI bars are drawn when available |

If kernel and DML plots appear to have different lower panels, check
`Xdistr`, `support.panel`, and `curve`. The support display is not estimator
specific: `"histogram"` draws many narrow bars, `"density"` draws a smoothed
filled density, `"rug"` draws ticks, and `"none"` suppresses support entirely.
When `curve = "both"`, support is embedded in each panel even if
`support.panel = "separate"` is requested. The two panels are drawn
side-by-side so the level and derivative curves remain readable. By default,
plot legends are drawn in a boxed area below the plot so they do not compete
with the curve. Use an explicit corner value such as `"topright"` only when
you intentionally want an inside-plot legend, or `legend.position = "none"` to
suppress it.

```{r curve-data, include=FALSE}
set.seed(101)
n_units_curve <- 60
curve_times <- 1:4
G_curve  <- runif(n_units_curve)
x_curve  <- rnorm(n_units_curve)

curve_panel <- data.frame(
  id   = rep(seq_len(n_units_curve), each = length(curve_times)),
  time = rep(curve_times, times = n_units_curve),
  G    = rep(G_curve, each = length(curve_times)),
  x1   = rep(x_curve, each = length(curve_times))
)

curve_panel$Y <- with(
  curve_panel,
  0.5 * x1 + 0.1 * time +
    ifelse(time >= 3, (time - 2) * (0.8 * G + 0.2 * G^2), 0) +
    rnorm(nrow(curve_panel), sd = 0.15)
)

prepared_curve <- fdid_prepare(
  data = curve_panel,
  Y_label = "Y",
  X_labels = "x1",
  G_label = "G",
  unit_label = "id",
  time_label = "time"
)

eval_curve <- seq(0.2, 0.8, length.out = 7)

fdid_results_kernel_curve <- fdid(
  s = prepared_curve,
  tr_period = 3:4,
  ref_period = 2,
  method = "kernel",
  eval_g = eval_curve,
  h0 = 0.25
)

fdid_results_kernel_boot_curve <- fdid(
  s = prepared_curve,
  tr_period = 3:4,
  ref_period = 2,
  method = "kernel",
  vartype = "bootstrap",
  eval_g = eval_curve,
  h0 = 0.25,
  boot = 19
)

fdid_results_flex_curve <- fdid(
  s = prepared_curve,
  tr_period = 3:4,
  ref_period = 2,
  method = "dml_flex",
  learner = "grf",
  K = 3,
  S = 1,
  eval_g = eval_curve
)

fdid_results_incremental_scalar <- fdid(
  s = prepared_curve,
  tr_period = 3:4,
  ref_period = 2,
  method = "dml_incremental",
  learner = "grf",
  K = 3,
  S = 1
)
```

The DML objects in this visualization chapter use `learner = "grf"` so the
plot examples match the main DML tutorial chapter.

```{r curve-kernel}
plot(fdid_results_kernel_curve, type = "curve", curve = "derivative",
     main = "Kernel Derivative Curve")
```

```{r curve-kernel-level}
plot(fdid_results_kernel_curve, type = "curve", curve = "level",
     main = "Kernel Level Curve")
```

```{r curve-flex}
plot(fdid_results_flex_curve, type = "curve", curve = "level",
     main = "DML Flexible Dose-Response Curve")
```

```{r curve-flex-derivative}
plot(fdid_results_flex_curve, type = "curve", curve = "derivative",
     main = "DML Flexible Derivative Curve")
summary(fdid_results_incremental_scalar)
```

### Curve, interval, and support display

`curve = "both"` draws level and derivative curves from the same fitted object.
This is useful when checking whether a visually smooth level curve implies a
plausible derivative curve.

```{r curve-both, fig.width=10, fig.height=4.8}
plot(fdid_results_kernel_curve, type = "curve", curve = "both",
     Xdistr = "density", main = "Kernel Level and Derivative")
```

The `interval` argument chooses which uncertainty layer is drawn. For kernel
curves, robust fits provide analytical pointwise normal intervals. Kernel
bootstrap fits additionally store bootstrap percentile intervals and finite-grid
quantile-envelope bands. For `dml_flex`, the label printed in the legend tells
whether the band is a practical signal-residual multiplier band or, for
`signal_map = "blp_spline"`, a Gaussian max-$t$ BLP-spline covariance band.

```{r curve-intervals, fig.width=9, fig.height=5.6}
old_par <- par(mfrow = c(1, 2))
plot(fdid_results_kernel_boot_curve, type = "curve", curve = "level",
     interval = "pointwise", Xdistr = "none",
     main = "Pointwise interval")
plot(fdid_results_kernel_boot_curve, type = "curve", curve = "level",
     interval = "both", Xdistr = "none",
     main = "Pointwise plus uniform band")
par(old_par)
```

The observed-support display can be made compact or more diagnostic depending
on the purpose of the figure.

```{r curve-support-displays, fig.width=9, fig.height=7}
old_par <- par(mfrow = c(2, 2))
plot(fdid_results_kernel_curve, type = "curve", curve = "level",
     Xdistr = "histogram", support.panel = "embedded",
     main = "Histogram support display")
plot(fdid_results_kernel_curve, type = "curve", curve = "level",
     Xdistr = "density", support.panel = "embedded",
     main = "Density support display")
plot(fdid_results_kernel_curve, type = "curve", curve = "level",
     Xdistr = "rug", support.panel = "embedded",
     main = "Rug support display")
plot(fdid_results_kernel_curve, type = "curve", curve = "level",
     Xdistr = "none", support.panel = "embedded",
     main = "No support display")
par(old_par)
```

When the support layer competes visually with the curve, move it to a separate
lower panel.

```{r curve-separate-support, fig.width=7, fig.height=5.2}
plot(fdid_results_flex_curve, type = "curve", curve = "level",
     Xdistr = "density", support.panel = "separate", show.eval_g = TRUE,
     main = "DML Curve With Separate Support Panel")
```

Use `diff.values` or `contrast.values` to mark substantively important values
of `G`. These arguments draw reference lines only; the numeric contrast should
still be reported with `fdid_contrast()`.

```{r curve-markers}
plot(fdid_results_flex_curve, type = "curve", curve = "level",
     Xdistr = "rug",
     contrast.values = c(0.3, 0.7),
     main = "Marked Contrast Values")
fdid_contrast(fdid_results_flex_curve, g0 = 0.3, g1 = 0.7)
```

### Fixed-reference contrast coefficient plot

A fixed-reference contrast coefficient plot fixes one reference value
$g_{\mathrm{ref}}$ and displays selected target contrasts,
$\theta(g) - \theta(g_{\mathrm{ref}})$ or
$\mu(g) - \mu(g_{\mathrm{ref}})$. Use
`type = "contrast"` for the built-in coefficient-style plot. The plot calls
`fdid_contrast()` for each displayed target value, so its intervals follow the
same inference rule as the reporting helper: stored curve covariance when
available, then replicate curves, then stored bands, then a labeled pointwise
fallback. The reference row is an algebraic zero anchor, not a separately
estimated contrast. Set `ci = FALSE` to show only point estimates.

```{r curve-reference-contrast}
plot(fdid_results_flex_curve, type = "contrast",
     target.type = "quantile",
     target.g = 5,
     main = "Five Quantile-Selected Contrasts")

plot(fdid_results_flex_curve, type = "contrast",
     ref.g = 0.5,
     target.g = c(0.2, 0.35, 0.65, 0.8),
     main = "Fixed-Reference Contrast Coefficients")
fdid_contrast(fdid_results_flex_curve, g0 = 0.5, g1 = 0.8)
```

## Comparison of methods

By default, the comparison plot uses a vertical layout (estimates stacked top to bottom). Set `vertical = FALSE` for a horizontal layout.

```{r}
plot(fdid_list(fdid_results_did, fdid_results_ols1, fdid_results_ols2,
               fdid_results_ebal, fdid_results_aipw), vertical = FALSE
     )
```

Common options:

-   `xlab` and `main` change labels/titles.

-   `ylim` sets the range of the estimates axis.

```{r}
plot(fdid_list(fdid_results_did, fdid_results_ols1, fdid_results_ols2,
               fdid_results_ebal, fdid_results_aipw),
     ylim = c(-5,5)
     )
```

```{r}
plot(fdid_list(fdid_results_did, fdid_results_ols1,
               fdid_results_ols2, fdid_results_ebal, fdid_results_aipw),
     xlab = 'Static Estimates',
     main = 'Comparing Different Methods', vertical = FALSE,
     ylim = c(-10,10)
     )
```

The labels reflect `method`, and add `vartype` when it is not the default `"robust"`. Other settings (e.g., `target.pop`) are not currently shown in the labels.

```{r}
plot(fdid_list(fdid_results_ols1, fdid_results_ols1_boot,
               fdid_results_ols1_jack, fdid_results_ols2, fdid_results_ols2_target_1,
               fdid_results_aipw),
     ylim = c(-5,1), vertical = TRUE
     )
```
