# Visualization Options {#sec-visualization}

```{r setup-visualization, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  message = FALSE,
  warning = FALSE
)
```

This chapter documents the current base-graphics plotting behavior. The
package-wide ggplot/autoplot modernization is deferred and should not be
promised as current behavior.

## Raw, Dynamic, And Overlap Displays

```{r plot-binary-displays, eval=FALSE}
plot(fit_ols1, type = "raw", connected = TRUE)
plot(fit_ols1, type = "dynamic", connected = TRUE)
plot(fit_aipw, type = "overlap")
```

Raw plots show descriptive outcome means. Dynamic plots show time-indexed FDID
estimates. Overlap plots are propensity-score diagnostics for binary methods
that store propensity scores.

## Continuous-G Curve Displays

```{r plot-curves, eval=FALSE}
plot(fit_kernel, type = "curve", curve = "level")
plot(fit_flex, type = "curve", curve = "level")
```

For `kernel` and `dml_flex`, the curve plot displays stored `eval_g` rows. It
does not re-estimate or densify the curve inside `plot()`.

## Level, Derivative, And Both

```{r plot-level-derivative, eval=FALSE}
plot(fit_kernel, type = "curve", curve = "level")
plot(fit_kernel, type = "curve", curve = "derivative")
plot(fit_kernel, type = "curve", curve = "both")
```

Use explicit `curve` values in reports so readers know which target is being
shown.

## Pointwise Versus Stored Uniform/Simultaneous Intervals

`interval = "pointwise"` displays pointwise intervals when available.
`interval = "uniform"` or `show.uniform.CI = TRUE` uses stored bands when the
fit contains them. Kernel bootstrap bands are quantile-envelope bands.
Non-BLP `dml_flex` bands are practical signal-mapping multiplier bands.
BLP-spline DML can use finite-grid covariance/Gaussian max-`t` bands when
usable.

## Support Displays: Rug, Density, Histogram, Separate Panel

```{r plot-support, eval=FALSE}
plot(fit_kernel, type = "curve", Xdistr = "rug")
plot(fit_kernel, type = "curve", Xdistr = "density")
plot(fit_kernel, type = "curve", Xdistr = "histogram",
     support.panel = "separate")
```

Support displays are descriptive overlays or panels. They help identify sparse
regions of `G`; they are not adjusted causal estimates.

## Evaluation-Grid And Marked-G Displays

```{r plot-markers, eval=FALSE}
plot(fit_kernel, type = "curve", show.eval_g = TRUE,
     diff.values = c(eval_g[2], eval_g[5], eval_g[8]))
```

Use evaluation-grid ticks and marked values to connect plot features with
reported contrasts or derivative values.

## Fixed-Reference Contrast Coefficient Plot

```{r plot-contrast, eval=FALSE}
plot(fit_kernel, type = "contrast", ref.g = "median",
     target.type = "quantile", target.g = 5)
```

The contrast plot displays level differences relative to a reference value. It
delegates estimates and intervals to `fdid_contrast()` and includes an
algebraic-zero reference row by default.

## Styling And Base-Graphics Arguments

Plots use base graphics and accept ordinary arguments such as `main`, `xlab`,
`ylab`, `xlim`, `ylim`, colors, line widths, point characters, and support
display controls. Save and restore `par()` if you change multi-panel layout in
your own scripts.

## What Will Change If The Package-Wide Plotting Layer Is Implemented

A future plotting-layer pass may add reusable plot-data builders and optional
ggplot/autoplot outputs. Until that happens, the stable public contract is the
current S3 `plot()` interface and its base-graphics output.
