# Reporting And Inference {#sec-reporting-inference}

```{r setup-reporting, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  message = FALSE,
  warning = FALSE
)
```

This chapter centralizes reporting and inference labels so the estimator
chapters can stay focused on workflows.

## `summary()`

```{r summary-reporting, eval=FALSE}
summary(fit_ols1)
summary(fit_kernel)
summary(fit_plr)
summary(fit_incremental)
```

`summary()` reports the scalar event and dynamic summaries stored by the fitted
object. For scalar DML methods, read the DML inference label rather than
assuming a bootstrap or refit procedure.

## `fdid_contrast()`

```{r contrast-reporting, eval=FALSE}
fdid_contrast(fit_kernel, g0 = eval_g[2], g1 = eval_g[7])
fdid_contrast(fit_flex, g0 = eval_g[2], g1 = eval_g[7],
              interpolate = TRUE)
```

`fdid_contrast()` reports a fixed-reference level contrast such as
`mu(g1) - mu(g0)` for kernel or `theta(g1) - theta(g0)` for `dml_flex`.

## `fdid_derivative()`

```{r derivative-reporting, eval=FALSE}
fdid_derivative(fit_kernel, g0 = eval_g[5])
fdid_derivative(fit_flex, g0 = eval_g[5])
```

`fdid_derivative()` reports a derivative value at a fixed `G`. It is distinct
from the scalar observed-population average derivative reported by
`dml_incremental`.

## When Contrast SEs Use Stored Covariance

Exact-grid contrasts can use stored covariance matrices when the fit provides
them. Kernel bootstrap fits may store curve covariance matrices. `dml_flex`
with `signal_map = "blp_spline"` can store finite-grid covariance matrices for
level and derivative curves.

## When Contrast SEs Use Replicate Curves

If stored replicate curves are available, contrast and value reporting can use
the empirical replicate distribution. Kernel bootstrap fits and DML multiplier
fits may provide these replicate arrays depending on options.

## When Intervals Are Pointwise Fallbacks

When covariance or replicate information is unavailable, reporting helpers may
fall back to pointwise endpoint information. Label these as pointwise or
approximate fallback intervals; do not present them as efficient contrast
inference.

## How To Interpret Simultaneous Bands

Simultaneous bands are finite-grid statements over the stored `eval_g` grid.
Kernel bootstrap bands are quantile-envelope bands. Non-BLP `dml_flex` bands
are practical signal-mapping multiplier bands. BLP-spline bands use
finite-grid covariance/Gaussian max-`t` calculations when available.

## What Is Bootstrap, What Is Analytical, What Is Multiplier

| Label | Meaning |
|---|---|
| Analytical | Standard errors computed from fitted analytical formulas |
| Bootstrap | Stored resampled curve replicates or percentile intervals |
| Multiplier | Multiplier draws over stored influence or signal values |
| Gaussian max-`t` | Finite-grid critical value from a covariance matrix |

The label should come from the fitted object whenever possible.

## Known Open Research-Level Inference Items

The package provides practical reporting for curves, contrasts, and
derivatives. Research-level work remains for theorem-level efficient uniform
DML bands and full influence-function storage for all non-BLP flexible curve
paths. Until that changes, use conservative labels in reports.
