# Sensitivity Analysis {#sec-sensitivity}

```{r setup-sensitivity, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  message = FALSE,
  warning = FALSE
)
```

Sensitivity checks should vary the assumptions that matter for the target. Do
not hide target changes inside one method-comparison table.

## Binary-G Sensitivity With sensemakr

For binary or discrete-`G` results, omitted-variable sensitivity can be studied
with `sensemakr` after constructing the appropriate regression object. Treat
this as a binary/discrete sensitivity analysis, not as evidence about the
continuous-`G` curve.

```{r sensemakr-example, eval=FALSE}
if (requireNamespace("sensemakr", quietly = TRUE)) {
  # Build the regression object for the binary-G target, then call sensemakr().
}
```

## Kernel Bandwidth And Support Sensitivity

Refit the kernel curve over plausible bandwidths and interior support windows.
Report whether the substantive pattern changes when tails or mass points are
excluded from the evaluation grid.

```{r kernel-bandwidth-sensitivity, eval=FALSE}
fit_kernel_h1 <- fdid(s_cont, tr_period, ref_period, method = "kernel",
                      eval_g = eval_g, h0 = 0.12)
fit_kernel_h2 <- fdid(s_cont, tr_period, ref_period, method = "kernel",
                      eval_g = eval_g, h0 = 0.24)
```

## Kernel Bootstrap Sensitivity

Use small bootstrap counts only for smoke checks. Increase bootstrap
replications for reports and verify that pointwise intervals and
quantile-envelope bands are stable enough for the claim being made.

## DML Learner Sensitivity

Compare transparent learners and flexible learners when optional packages are
available. Keep the target fixed while varying `learner`.

```{r dml-learner-sensitivity, eval=FALSE}
fit_flex_linear <- fdid(s_cont, tr_period, ref_period, method = "dml_flex",
                        learner = "linear", eval_g = eval_g)
fit_flex_ridge <- fdid(s_cont, tr_period, ref_period, method = "dml_flex",
                       learner = "ridge", eval_g = eval_g)
```

## DML Density-Method Sensitivity

For `dml_flex` and `dml_incremental`, vary `density_method` because the
conditional density enters the score. Large changes across density methods are
diagnostic, especially for the incremental average derivative.

## Cross-Fitting And Repeated-Split Sensitivity

Vary `K` and `S` to check whether fold assignment and repeated-split
aggregation drive the result. Report the values used in final tables.

## High-Dimensional Expansion Sensitivity

`fdid_expand_covariates()` can enrich the covariate design before DML
estimation. Treat this as a design sensitivity check: more features can reduce
misspecification but may increase variance and nuisance instability.

## How To Report Sensitivity Without Overclaiming

Report sensitivity by target. For example, say the kernel level curve is
stable across bandwidths, or the incremental average derivative changes under
different density methods. Avoid broad claims such as "all continuous-G
methods agree" unless the compared objects truly estimate the same target.
