# Continuous G {#sec-continuous}

```{r setup-cont, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 7,
  fig.height = 4.8
)
set.seed(20260530)
```

This section gives a guided workflow for continuous-$G$ FDID. The key point is
that continuous-$G$ analysis is **estimand-first**: choose the object you want
to report, then choose the method. A single continuous variable can support
several different empirical questions:

- a scalar slope, which summarizes the relationship with one number;
- a level curve, which reports the event-period effect at different values of
  $G$;
- a level contrast, which compares two meaningful values along the curve;
- a derivative value or derivative curve, which describes local marginal
  effects;
- an observed-population average derivative, which is a scalar marginal-effect
  estimand over the observed distribution of $G$.

These objects should not be mixed together. In particular, the scalar average
derivative from `method = "dml_incremental"` is not the average of the
`dml_flex` derivative curve.

| User question | Start with | Report |
|---|---|---|
| Is there one average relationship between $G$ and the event-period change? | `"ols1"`, `"ols2"`, or `"dml_plr"` | scalar slope |
| Does the event effect vary nonlinearly with $G$? | `"kernel"` or `"dml_flex"` | level curve |
| How different are two meaningful $G$ values? | `fdid_contrast()` | level contrast |
| What is the local marginal effect at one $G$ value? | `fdid_derivative()` | derivative value, interpreted cautiously |
| What is one average marginal effect over observed $G$? | `"dml_incremental"` | scalar observed average derivative |

::: {.callout-tip appearance="simple"}
Start with a scalar benchmark and a level curve. Add derivative and DML
sensitivity checks only when the substantive question requires them.
:::

## Data Setup And Support

```{r library-cont}
library(fdid)
data(fdid)
```

We use `lnpczupu` as the continuous baseline factor. It is the log-transformed
genealogy-book density, `log(pczupu + 1)`. As in the binary-$G$ chapter, the
event window is the Great Famine period and the reference period is 1957.

```{r data-cont}
mortality$uniqueid <- paste(as.character(mortality$provid),
                            as.character(mortality$countyid), sep = "-")

covar <- c("avggrain", "nograin", "urban", "dis_bj", "dis_pc",
           "rice", "minority", "edu", "lnpop")
tr_period <- 1958:1961
ref_period <- 1957

s_cont <- fdid_prepare(
  data = transform(mortality, G_cont = lnpczupu),
  Y_label = "mortality",
  X_labels = covar,
  G_label = "G_cont",
  unit_label = "uniqueid",
  time_label = "year"
)
```

For curve methods, `eval_g` is the grid where the curve is estimated. It is not
only a plotting option. Choose grid values where the data have support. The
mortality data have many counties with `lnpczupu = 0`, so this example reports
curves over the interior of the positive support.

There are two separate choices here:

- the **support range**: which part of the observed $G$ distribution is credible
  enough to report;
- the **grid density**: how many points to estimate inside that range.

The examples below use a short grid so the tutorial renders quickly. In a final
analysis, use a denser grid over the same support range.

```{r g-support}
g_positive <- s_cont$G[s_cont$G > 0]
eval_g <- seq(
  stats::quantile(g_positive, 0.15, na.rm = TRUE),
  stats::quantile(g_positive, 0.85, na.rm = TRUE),
  length.out = 11
)

support_table <- data.frame(
  quantity = c("n units", "share G = 0", "curve grid lower", "curve grid upper"),
  value = c(nrow(s_cont), mean(s_cont$G == 0), min(eval_g), max(eval_g))
)
knitr::kable(support_table, digits = 4)
```

```{r g-support-plot}
hist(s_cont$G, breaks = 35, col = "gray85", border = "white",
     xlab = "lnpczupu", main = "Support of continuous G")
abline(v = range(eval_g), col = c("steelblue", "firebrick"), lwd = 2)
rug(s_cont$G, col = grDevices::adjustcolor("black", 0.25))
legend("topright", legend = "curve grid boundary",
       col = "steelblue", lwd = 2, bty = "n")
```

The vertical lines mark the curve grid boundary. The mass at zero is visible in
the histogram. We do not turn continuous $G$ into bins; we simply avoid asking a
curve estimator to extrapolate into poorly supported regions.

## Scalar Continuous-G Benchmarks

Scalar methods answer whether the event-period outcome change has one average
relationship with continuous $G$. They are useful benchmarks, but they do not
describe nonlinear dose-response shapes.

The three scalar benchmarks below differ in how they adjust for observed
covariates:

- `ols1` uses additive covariate adjustment;
- `ols2` adds $G \times X$ interactions and then averages over the target
  population;
- `dml_plr` uses a partially linear DML target with cross-fitted nuisance
  functions.

If these scalar estimates point in very different directions, inspect the
support of $G$, covariate balance, and nonlinear curve estimates before
reporting a single scalar result.

```{r scalar-benchmarks}
event_row <- function(x, method, target) {
  ev <- x$est$event
  data.frame(
    method = method,
    target = target,
    estimate = as.numeric(ev$Estimate[1]),
    std_error = as.numeric(ev$Std.Error[1]),
    ci_lower = as.numeric(ev$CI_Lower[1]),
    ci_upper = as.numeric(ev$CI_Upper[1])
  )
}

fit_ols1_cont <- fdid(
  s = s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "ols1"
)

fit_ols2_cont <- fdid(
  s = s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "ols2"
)

fit_dml_plr <- fdid(
  s = s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "dml_plr",
  learner = "linear",
  K = 3,
  S = 1
)

scalar_table <- rbind(
  event_row(fit_ols1_cont, "ols1", "additive adjusted slope"),
  event_row(fit_ols2_cont, "ols2", "interaction adjusted slope"),
  event_row(fit_dml_plr, "dml_plr", "partially linear DML slope")
)

knitr::kable(scalar_table, digits = 4)
```

Interpret these as one-number summaries of the event-period relationship with
`lnpczupu`. They do not say whether the effect is steeper at low social capital
than at high social capital, or whether the relationship changes sign.

## Level Curves

Use a curve method when the main question is how the event-period FDID effect
varies over the support of continuous $G$.

`method = "kernel"` is the most transparent first curve method. It locally
smooths the event-period outcome change around each value of `eval_g`.

```{r kernel-curve}
fit_kernel <- fdid(
  s = s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "kernel",
  eval_g = eval_g,
  vartype = "bootstrap",
  boot = 5
)

plot(fit_kernel, type = "curve", curve = "level",
     main = "Kernel level curve")
```

The kernel level curve is a useful first diagnostic because it keeps the
continuous-$G$ target visually explicit. Its main limitation is that it is not a
high-dimensional adjustment strategy; if confounding adjustment is central, use
it as a benchmark rather than the final result.

`method = "dml_flex"` targets an adjusted nonlinear level curve,
$\theta(g) = E_X[\mu(g, X)]$. It uses cross-fitting, nuisance models, and a
signal-to-$G$ map. Start with simple choices, then vary them as sensitivity
checks.

The important tuning choices are:

- `learner`: the nuisance-function learner;
- `density_method`: how the conditional density of $G$ given covariates is
  estimated;
- `signal_map`: how the orthogonal signal is mapped back to values of $G$;
- `K` and `S`: the number of folds and repeated sample splits.

The tutorial uses `learner = "linear"`, `K = 3`, and `S = 1` to keep the
example easy to render. For a final analysis, consider more splits and
sensitivity checks.

```{r dml-flex-curve}
fit_flex <- fdid(
  s = s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "dml_flex",
  learner = "linear",
  K = 3,
  S = 1,
  eval_g = eval_g,
  signal_map = "spline",
  density_method = "residual_kde",
  boot = 5
)

plot(fit_flex, type = "curve", curve = "level",
     main = "DML flexible level curve")
```

The DML level curve is the main adjusted nonlinear object. Its scalar
`$est$event` component is only a compatibility summary; the empirical object to
inspect is the stored curve over `eval_g`.

```{r dml-flex-curve-table}
flex_curve_table <- data.frame(
  G = fit_flex$eval_g,
  theta_hat = fit_flex$curve_event$theta_hat,
  se_theta = fit_flex$curve_event$se_theta
)

knitr::kable(head(flex_curve_table, 6), digits = 4)
```

## Reporting Contrasts And Derivatives

A full curve is useful visually, but a level contrast is often easier to report
in text. `fdid_contrast()` compares two values along the stored fitted curve.

Here we compare the upper end of the supported grid to the lower end. This is a
continuous-curve contrast, not a binary high-versus-low treatment estimator.

```{r contrast-report}
fdid_contrast(
  fit_flex,
  g0 = min(eval_g),
  g1 = max(eval_g)
)
```

A typical writeup should name the two values being compared and state the
support rule used to choose them. For example: "We compare the fitted curve at
the 85th percentile of positive `lnpczupu` with the fitted curve at the 15th
percentile of positive `lnpczupu`."

Derivative curves and derivative values answer sharper local questions. They
are more sensitive than level curves and should usually be treated as secondary
evidence.

```{r derivative-report}
fdid_derivative(
  fit_flex,
  g0 = stats::median(eval_g)
)
```

For a visual derivative diagnostic:

```{r derivative-plot}
plot(fit_flex, type = "curve", curve = "derivative",
     main = "DML flexible derivative curve")
```

Derivative results are most defensible when the level curve is smooth and the
substantive question is explicitly local. If the level curve has kinks, flat
segments, or wide uncertainty bands, lead with level contrasts instead.

## Scalar Average Derivative

`method = "dml_incremental"` targets one observed-population average derivative,
$E[\partial_g \mu(G, X)]$. It is not the same as averaging the derivative curve
from `dml_flex`.

Because the estimator depends on the derivative of the outcome regression and
the derivative of the log conditional density of $G$ given $X$, the density
method is a real sensitivity choice. The example reports the result as a plain
table rather than printing `summary()`, because the console summary uses
box-drawing separators that can render as escaped Unicode under some R/Quarto
locales.

```{r incremental}
fit_incremental_resid <- fdid(
  s = s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "dml_incremental",
  learner = "linear",
  K = 3,
  S = 1,
  density_method = "residual_kde"
)

fit_incremental_locscale <- fdid(
  s = s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "dml_incremental",
  learner = "linear",
  K = 3,
  S = 1,
  density_method = "location_scale"
)

incremental_row <- function(x, density_method) {
  ev <- x$est$event
  data.frame(
    method = x$method,
    learner = x$learner,
    density_method = density_method,
    estimate = as.numeric(ev$Estimate[1]),
    std_error = as.numeric(ev$Std.Error[1]),
    ci_lower = as.numeric(ev$CI_Lower[1]),
    ci_upper = as.numeric(ev$CI_Upper[1])
  )
}

incremental_table <- rbind(
  incremental_row(fit_incremental_resid, "residual_kde"),
  incremental_row(fit_incremental_locscale, "location_scale")
)

knitr::kable(incremental_table, digits = 4)
```

Use `dml_incremental` when the target is one average marginal effect over the
observed distribution of $G$, not a curve.

## Sensitivity Checks

The main empirical story should not cross every option with every other option.
Instead, vary one dimension at a time:

1. raw covariates versus expanded covariates;
2. simple versus flexible learners;
3. `density_method` for `dml_flex` and `dml_incremental`;
4. `signal_map` for `dml_flex`.

For richer nuisance adjustment, `fdid_expand_covariates()` can expand the
prepared covariates into polynomial or spline bases.

```{r hd-expansion-cont}
s_cont_hd <- fdid_expand_covariates(
  s_cont,
  basis_type = "polynomial",
  poly_degree = 2,
  include_interactions = TRUE
)

hd_table <- data.frame(
  design = c("raw covariates", "polynomial plus interactions"),
  covariate_columns = c(
    length(grep("^x[0-9]+$", names(s_cont))),
    length(grep("^x[0-9]+$", names(s_cont_hd)))
  )
)

knitr::kable(hd_table)
```

The expanded design can be useful with regularized or flexible learners, but it
also makes the nuisance problem harder. Treat it as a robustness check, not an
automatic improvement.

```{r learner-availability-cont}
learner_package <- c(
  linear = NA_character_,
  gam = "mgcv",
  lasso = "glmnet",
  ridge = "glmnet",
  elasticnet = "glmnet",
  ranger = "ranger",
  grf = "grf",
  nnet = "nnet",
  xgboost = "xgboost"
)

learner_table <- data.frame(
  learner = names(learner_package),
  package = unname(learner_package),
  available = vapply(learner_package, function(pkg) {
    is.na(pkg) || requireNamespace(pkg, quietly = TRUE)
  }, logical(1))
)

knitr::kable(learner_table)
```

Use optional learners only when their backend packages are installed and when
the additional flexibility is part of the analysis plan.

## Practical Checklist

For applied continuous-$G$ reporting, include:

- the baseline factor used as $G$;
- the event window and reference period;
- the support range used for `eval_g`;
- whether the main result is a scalar slope, level curve, level contrast,
  derivative value, or average derivative;
- the method and key tuning choices, especially `learner`, `density_method`,
  and `signal_map` for DML curves.

The practical sequence is: scalar benchmark, level curve, level contrast, and
then derivative or DML sensitivity checks only when those sharper claims matter.
