# Visualization Options {#sec-visualization}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)
```

In this chapter, we explore various visualization options available in the `fdid` package.

```{r}
library(fdid)
```

```{r readcsv, include=FALSE}
data(fdid)
```

```{r gen, include=FALSE}
mortality$uniqueid <- paste(as.character(mortality$provid), as.character(mortality$countyid), sep = "-")
mortality$highzupu <- ifelse(mortality$pczupu >= median(mortality$pczupu, na.rm = TRUE), 1, 0)
```

```{r parameters, include=FALSE}
tr_period <- 1958:1961
ref_period <- 1957
treat <- "highzupu"
Y <- "mortality"
time <- "year"
unit <- "uniqueid"
covar <- c("avggrain", "nograin", "urban", "dis_bj", "dis_pc",
           "rice", "minority", "edu", "lnpop")
```

```{r prepare, include=FALSE}
prepared_d <- fdid_prepare(
  data = mortality,
  Y_label = Y,
  X_labels = covar,
  G_label = treat,
  unit_label = unit,
  time_label = time
)
```

```{r ols1, include=FALSE}
fdid_results_ols1 <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  vartype = "robust",
)
```

```{r ols2, include=FALSE}
fdid_results_ols2 <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "ols2",
  vartype = "robust",
)
```

```{r did, include=FALSE}
fdid_results_did <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "did",
  vartype = "robust",
)
```

```{r ebal, cache=TRUE, include=FALSE}
fdid_results_ebal <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "ebal",
  vartype = "robust",
  target.pop = "1"
)
```

```{r ipw, cache=TRUE, include=FALSE}
fdid_results_ipw <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "ipw",
  vartype = "robust",
)
```

```{r aipw, cache=TRUE, include=FALSE}
fdid_results_aipw <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "aipw",
  vartype = "robust",
)
```

```{r boot, cache=TRUE, include=FALSE}
fdid_results_ols1_boot <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  vartype = "bootstrap",
  parallel = TRUE,
  cores = 6,
  nsims = 100
)
```

```{r jackknife, cache=TRUE, include=FALSE}
fdid_results_ols1_jack <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  vartype = "jackknife",
  parallel = TRUE,
  cores = 6,
)
```

```{r ols2_target1, cache=TRUE, include=FALSE}
fdid_results_ols2_target_1 <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "ols2",
  vartype = "robust",
  target.pop = "1"
)
```


## Raw means

-  `ci = FALSE` removes the 95% confidence intervals.

```{r}
plot(fdid_results_ols1, type = "raw", ci = FALSE)
```

-  `connected = TRUE` connects points with lines (default: `FALSE`).

```{r}
plot(fdid_results_ols1, type = "raw", connected = TRUE)
```

-  `group_labels` renames the two groups.

```{r}
plot(fdid_results_ols1, type = "raw", connected = TRUE, 
     group_labels = c("Low Social Capital", "High Social Capital"))
```

-   `palette` chooses an RColorBrewer palette (default: "Set2").

```{r}
plot(fdid_results_ols1, type = "raw", connected = TRUE, 
     group_labels = c("Low Social Capital", "High Social Capital"), palette = "Set1")
```

-   `shade_periods` controls the position of shaded region. Default is `x$tr_period`, i.e. the event window.     
    - By setting `shade_periods = c(1954, 1957)`, the pre-event periods can be set shaded.
    - By setting `shade_periods = NULL`, the user can also remove the shaded region.

```{r}
plot(fdid_results_ols1, type = "raw", connected = TRUE, 
     group_labels = c("Low Social Capital", "High Social Capital"),
     shade_periods = c(1954, 1957))
```

```{r}
plot(fdid_results_ols1, type = "raw", connected = TRUE, 
     group_labels = c("Low Social Capital", "High Social Capital"),
     shade_periods = NULL)
```

-   `alpha_shade` controls the transparency of the shaded region.

```{r}
plot(fdid_results_ols1, type = "raw", connected = TRUE, 
     group_labels = c("Low Social Capital", "High Social Capital"),
     alpha_shade = 0.6)
```

-   The user can also set `ylim`, `xlab`, `ylab`, and `main` (currently, xlim is not supported).

```{r}
plot(fdid_results_ols1, type = "raw", connected = TRUE, 
     group_labels = c("Low Social Capital", "High Social Capital"), 
     ylim = c(0, 80), xlab = "Year", ylab = "Mortality", main = "Raw Mortality")
```

## Dynamic estimates

-   `ci = FALSE` removes the 95% confidence intervals.

```{r}
plot(fdid_results_ols1, type = "dynamic", ci = FALSE)
```

-   `connected = TRUE` connects points with lines (default: `FALSE`).

```{r}
plot(fdid_results_ols1, type = "dynamic", connected = TRUE)
```

-   `shade_periods`, `alpha_shade`, `ylim`, `xlab`, `ylab`, and `main` customize the plot as above.

```{r}
plot(fdid_results_ols1, type = "dynamic", connected = TRUE,
     shade_periods = c(1954, 1957),
     alpha_shade = 0.8, ylim = c(-10, 4),
     ylab = "Estimates", xlab = "Year", 
     main = "Effect Modification of High Social Capital"
     )
```

-   `xlim` zooms into a subset of years.

```{r}
plot(fdid_results_ols1, type = "dynamic",
     alpha_shade = 0.8, ylim = c(-10, 4), xlim = c(1954, 1961),
     ylab = "Estimates", xlab = "Year", 
     main = "Hiding Post-Event Periods"
     )
```

## Overlap test 

The user can adjust `palette`, `ylim`, `xlab`, and `main`:

```{r}
plot(fdid_results_aipw, type = "overlap",
     palette = "Set1", ylim = c(-0.1, 0.1),
     xlab = 'Score', main = 'Diagnostic Test for Overlap Assumption'
     )
```

## Continuous-G curves

For `method = "kernel"` and `method = "dml_flex"`, `type = "curve"` plots the
event-period curve over the evaluation grid for $G$.

The evaluation grid is part of estimation. If `eval_g` is supplied, `kernel`
and `dml_flex` estimate curve values at exactly those grid points; if it is
left as `NULL`, the package builds its default grid. The curve plot displays
the stored `eval_g` rows and does not estimate additional points while
plotting.

Use the `curve` argument to choose the estimand:

| Method | Level curve | Derivative curve |
|---|---|---|
| `"kernel"` | `curve = "level"` | `curve = "derivative"` |
| `"dml_flex"` | `curve = "level"` | `curve = "derivative"` |

`method = "dml_incremental"` is scalar; it reports the observed-population
average derivative through `summary()` rather than `type = "curve"`.

For numeric reporting from these stored curves, use `fdid_contrast()` for
level comparisons and `fdid_derivative()` for fixed-$g$ derivative values.

```{r curve-data, include=FALSE}
set.seed(101)
n_units_curve <- 60
curve_times <- 1:4
G_curve  <- runif(n_units_curve)
x_curve  <- rnorm(n_units_curve)

curve_panel <- data.frame(
  id   = rep(seq_len(n_units_curve), each = length(curve_times)),
  time = rep(curve_times, times = n_units_curve),
  G    = rep(G_curve, each = length(curve_times)),
  x1   = rep(x_curve, each = length(curve_times))
)

curve_panel$Y <- with(
  curve_panel,
  0.5 * x1 + 0.1 * time +
    ifelse(time >= 3, (time - 2) * (0.8 * G + 0.2 * G^2), 0) +
    rnorm(nrow(curve_panel), sd = 0.15)
)

prepared_curve <- fdid_prepare(
  data = curve_panel,
  Y_label = "Y",
  X_labels = "x1",
  G_label = "G",
  unit_label = "id",
  time_label = "time"
)

eval_curve <- seq(0.2, 0.8, length.out = 7)

fdid_results_kernel_curve <- fdid(
  s = prepared_curve,
  tr_period = 3:4,
  ref_period = 2,
  method = "kernel",
  eval_g = eval_curve,
  h0 = 0.25
)

fdid_results_flex_curve <- fdid(
  s = prepared_curve,
  tr_period = 3:4,
  ref_period = 2,
  method = "dml_flex",
  learner = "linear",
  K = 3,
  S = 1,
  eval_g = eval_curve
)

fdid_results_incremental_scalar <- fdid(
  s = prepared_curve,
  tr_period = 3:4,
  ref_period = 2,
  method = "dml_incremental",
  learner = "linear",
  K = 3,
  S = 1
)
```

```{r curve-kernel}
plot(fdid_results_kernel_curve, type = "curve", curve = "derivative",
     main = "Kernel Derivative Curve")
```

```{r curve-kernel-level}
plot(fdid_results_kernel_curve, type = "curve", curve = "level",
     main = "Kernel Level Curve")
```

```{r curve-flex}
plot(fdid_results_flex_curve, type = "curve", curve = "level",
     main = "DML Flexible Dose-Response Curve")
```

```{r curve-flex-derivative}
plot(fdid_results_flex_curve, type = "curve", curve = "derivative",
     main = "DML Flexible Derivative Curve")
summary(fdid_results_incremental_scalar)
```

## Comparison of methods

By default, the comparison plot uses a vertical layout (estimates stacked top to bottom). Set `vertical = FALSE` for a horizontal layout.

```{r}
plot(fdid_list(fdid_results_did, fdid_results_ols1, fdid_results_ols2, 
               fdid_results_ebal, fdid_results_aipw), vertical = FALSE
     )
```

Common options:

-   `xlab` and `main` change labels/titles.

-   `ylim` sets the range of the estimates axis.

```{r}
plot(fdid_list(fdid_results_did, fdid_results_ols1, fdid_results_ols2, 
               fdid_results_ebal, fdid_results_aipw),
     ylim = c(-5,5)
     )
```

```{r}
plot(fdid_list(fdid_results_did, fdid_results_ols1, 
               fdid_results_ols2, fdid_results_ebal, fdid_results_aipw),
     xlab = 'Static Estimates',
     main = 'Comparing Different Methods', vertical = FALSE,
     ylim = c(-10,10)
     )
```

The labels reflect `method`, and add `vartype` when it is not the default `"robust"`. Other settings (e.g., `target.pop`) are not currently shown in the labels.

```{r}
plot(fdid_list(fdid_results_ols1, fdid_results_ols1_boot,
               fdid_results_ols1_jack, fdid_results_ols2, fdid_results_ols2_target_1,
               fdid_results_aipw),
     ylim = c(-5,1), vertical = TRUE
     )
```
