# Continuous G {#sec-continuous}

```{r setup-cont, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 7,
  fig.height = 4.8
)
set.seed(20260516)

fit_fdid <- function(...) {
  tryCatch(
    fdid(...),
    error = function(e) {
      structure(list(message = conditionMessage(e)), class = "fdid_fit_error")
    }
  )
}

is_fit_error <- function(x) inherits(x, "fdid_fit_error")

event_row <- function(obj, method, target, learner = NA_character_,
                      design = NA_character_, density_method = NA_character_,
                      signal_map = NA_character_) {
  if (is_fit_error(obj)) {
    return(data.frame(
      method = method, target = target, learner = learner, design = design,
      density_method = density_method, signal_map = signal_map,
      estimate = NA_real_, std_error = NA_real_, ci_lower = NA_real_,
      ci_upper = NA_real_, note = obj$message
    ))
  }
  ev <- obj$est$event
  data.frame(
    method = method,
    target = target,
    learner = learner,
    design = design,
    density_method = density_method,
    signal_map = signal_map,
    estimate = as.numeric(ev$Estimate[1]),
    std_error = as.numeric(ev$Std.Error[1]),
    ci_lower = as.numeric(ev$CI_Lower[1]),
    ci_upper = as.numeric(ev$CI_Upper[1]),
    note = ""
  )
}

curve_row <- function(obj, label, curve = c("level", "derivative")) {
  curve <- match.arg(curve)
  if (is_fit_error(obj) || is.null(obj$curve_event)) {
    return(data.frame())
  }
  if (curve == "level") {
    y <- obj$curve_event$theta_hat
    if (is.null(y)) y <- obj$curve_event$mu_hat
  } else {
    y <- obj$curve_event$delta_hat
  }
  data.frame(label = label, G = obj$eval_g, estimate = as.numeric(y))
}

plot_curve_overlay <- function(curves, truth = NULL, ylab = "Estimate",
                               main = NULL) {
  if (nrow(curves) == 0L) {
    plot.new()
    title(main = main)
    text(0.5, 0.5, "No fitted curves available")
    return(invisible(NULL))
  }
  labs <- unique(curves$label)
  cols <- grDevices::hcl.colors(length(labs), "Dark 3")
  ylim <- range(curves$estimate, if (!is.null(truth)) truth$estimate, na.rm = TRUE)
  plot(range(curves$G), ylim, type = "n",
       xlab = "Baseline factor G", ylab = ylab, main = main)
  if (!is.null(truth)) {
    lines(truth$G, truth$estimate, lwd = 2.5, lty = 2, col = "gray20")
  }
  for (j in seq_along(labs)) {
    cc <- curves[curves$label == labs[j], ]
    lines(cc$G, cc$estimate, lwd = 2, col = cols[j])
  }
  legend("topright",
         legend = c(if (!is.null(truth)) "truth" else character(0), labs),
         col = c(if (!is.null(truth)) "gray20" else character(0), cols),
         lty = c(if (!is.null(truth)) 2 else integer(0), rep(1, length(labs))),
         lwd = c(if (!is.null(truth)) 2.5 else numeric(0), rep(2, length(labs))),
         bty = "n", cex = 0.8)
  invisible(NULL)
}

learner_package <- c(
  linear = NA_character_,
  gam = "mgcv",
  lasso = "glmnet",
  ridge = "glmnet",
  elasticnet = "glmnet",
  ranger = "ranger",
  grf = "grf",
  nnet = "nnet",
  xgboost = "xgboost"
)

learner_available <- function(learner) {
  pkg <- learner_package[[learner]]
  is.na(pkg) || requireNamespace(pkg, quietly = TRUE)
}
```

This chapter reorganizes the continuous-treatment workflow around the estimand
first. The question is not only "which method should I run?" but "which
continuous-$G$ object do I want to report?"

The chapter uses two tracks:

- A mortality running example, which is compact and interpretable.
- Known-truth synthetic examples, which show what the estimators can and cannot
  recover under controlled designs.

## Local RStudio Workflow

When rendering from RStudio, make sure RStudio loads the local development
package, not an older installed CRAN or GitHub copy.

```{r local-install-example, eval=FALSE}
remove.packages("fdid")
devtools::install_local(
  "/Users/liuenhan/Desktop/Statsclaw-FDID/fdid-main",
  force = TRUE,
  upgrade = "never"
)

.rs.restartR()
library(fdid)
packageVersion("fdid")
normalizePath(find.package("fdid"))
```

The expected package path should point inside the local R library that received
the install, and the installed source should match
`/Users/liuenhan/Desktop/Statsclaw-FDID/fdid-main`. Then open the tutorial
project in RStudio and click **Render**. From Terminal, RStudio's bundled Quarto
can also render the book:

```{bash local-quarto-render, eval=FALSE}
/Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto render tutorial
```

## Estimator Map

All continuous-$G$ methods enter through `fdid(..., method = ...)`.

| User question | Target | Method | Main output | Main caveat |
|---|---|---|---|---|
| Is there one average slope? | linear slope | `"did"`, `"ols1"`, `"ols2"` | scalar table | misspecifies nonlinear curves |
| Is there one ML-adjusted slope? | partially linear slope | `"dml_plr"` | scalar table | PLR target only |
| What is the visual dose-response? | level and derivative curves | `"kernel"` | curve plots | weak with high-dimensional confounding |
| What is the nonlinear adjusted curve? | `theta(g)` and `theta_prime(g)` | `"dml_flex"` | level and derivative plots | derivative is sensitive |
| What is the observed average derivative? | `E[partial_g mu(G, X)]` | `"dml_incremental"` | scalar table | density-score sensitive |
| What if I dichotomize `G`? | binary AIPW/IRM effect | `"dml_binary"` | scalar table | binary estimand only |

The curve plot API makes the target explicit:

```{r curve-api-example, eval=FALSE}
plot(result_kernel, type = "curve", curve = "level")
plot(result_kernel, type = "curve", curve = "derivative")

plot(result_flex, type = "curve", curve = "level")
plot(result_flex, type = "curve", curve = "derivative")
```

`method = "dml_incremental"` is scalar. It does not have a curve plot.

## Data Setup And Support

```{r library-cont}
library(fdid)
data(fdid)
```

```{r data-cont}
mortality$uniqueid <- paste(as.character(mortality$provid),
                            as.character(mortality$countyid), sep = "-")

covar <- c("avggrain", "nograin", "urban", "dis_bj", "dis_pc",
           "rice", "minority", "edu", "lnpop")
tr_period <- 1958:1961
ref_period <- 1957

s_cont <- fdid_prepare(
  data = transform(mortality, G_cont = lnpczupu),
  Y_label = "mortality",
  X_labels = covar,
  G_label = "G_cont",
  unit_label = "uniqueid",
  time_label = "year"
)
```

The mortality `lnpczupu` variable has a large mass at zero. The main curve grid
therefore uses the interior of the positive support. This is a support choice,
not a binning estimator: the estimator still smooths over continuous `G`.

For `method = "kernel"` and `method = "dml_flex"`, the `eval_g` argument is
the estimation grid, not only a plotting grid. The package estimates and stores
curve quantities at those values. In the rendered tutorial below,
`eval_g_curve` is kept shorter so the examples run quickly; in a substantive
analysis, use the default 50-point grid or supply a denser grid over the
support you want to report. The separate `g_landmarks` values are only
landmarks for labels and interpretation.

```{r g-support}
g_positive <- s_cont$G[s_cont$G > 0]
eval_g_main <- seq(
  stats::quantile(g_positive, 0.15, na.rm = TRUE),
  stats::quantile(g_positive, 0.85, na.rm = TRUE),
  length.out = 25
)
eval_g_curve <- eval_g_main[round(seq(1, length(eval_g_main), length.out = 13))]
g_landmarks <- stats::quantile(g_positive, c(0.15, 0.25, 0.50, 0.75, 0.85),
                               na.rm = TRUE)

support_table <- data.frame(
  quantity = c("n units", "share G = 0", "positive-support 15%", "positive-support 85%"),
  value = c(nrow(s_cont), mean(s_cont$G == 0), eval_g_main[1],
            eval_g_main[length(eval_g_main)])
)
knitr::kable(support_table, digits = 4)
```

```{r g-support-plot}
hist(s_cont$G, breaks = 35, col = "gray85", border = "white",
     xlab = "lnpczupu", main = "Support of continuous G")
abline(v = range(eval_g_main), col = c("steelblue", "firebrick"), lwd = 2)
abline(v = g_landmarks, col = "gray35", lty = 3)
rug(s_cont$G, col = grDevices::adjustcolor("black", 0.25))
legend("topright",
       legend = c("curve grid boundary", "representative quantiles"),
       col = c("steelblue", "gray35"), lty = c(1, 3), lwd = c(2, 1),
       bty = "n")
```

The five landmark values are only used for labels and concise interpretation.
They are not five treatment bins, and they do not control the curve estimation
unless they are explicitly passed as `eval_g`.

## Scalar Benchmarks On Mortality

These methods answer whether the event-period outcome change has one average
relationship with continuous `G`. They do not settle whether the relationship is
nonlinear.

```{r mortality-scalar-benchmarks}
fit_scalar_benchmarks <- list(
  did = fit_fdid(s_cont, tr_period = tr_period, ref_period = ref_period,
                 method = "did"),
  ols1 = fit_fdid(s_cont, tr_period = tr_period, ref_period = ref_period,
                  method = "ols1"),
  ols2 = fit_fdid(s_cont, tr_period = tr_period, ref_period = ref_period,
                  method = "ols2"),
  dml_plr = fit_fdid(s_cont, tr_period = tr_period, ref_period = ref_period,
                     method = "dml_plr", learner = "linear", K = 3, S = 1)
)

scalar_table <- do.call(rbind, list(
  event_row(fit_scalar_benchmarks$did, "did", "unadjusted linear slope"),
  event_row(fit_scalar_benchmarks$ols1, "ols1", "additive adjusted slope"),
  event_row(fit_scalar_benchmarks$ols2, "ols2", "interaction adjusted slope"),
  event_row(fit_scalar_benchmarks$dml_plr, "dml_plr", "partially linear DML slope",
            learner = "linear")
))

knitr::kable(scalar_table[, c("method", "target", "estimate", "std_error",
                              "ci_lower", "ci_upper", "note")],
             digits = 4)
```

## Curve Estimators On Mortality

### Kernel Level And Derivative Curves

`method = "kernel"` locally smooths the event-period outcome change around each
evaluation point `g`. It reports a level curve `mu_hat(g)` and a derivative
curve `delta_hat(g)`.

The fitted curve has one row per value in `eval_g`. `plot(type = "curve")`
draws those stored estimates and their intervals; it does not add new
estimation points after the fit.

```{r mortality-kernel}
fit_kernel_main <- fit_fdid(
  s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "kernel",
  eval_g = eval_g_curve,
  K_folds = 3,
  vartype = "bootstrap",
  boot = 3
)

if (!is_fit_error(fit_kernel_main)) {
  par(mfrow = c(1, 2))
  plot(fit_kernel_main, type = "curve", curve = "level",
       main = "Kernel level curve")
  plot(fit_kernel_main, type = "curve", curve = "derivative",
       main = "Kernel derivative curve")
  par(mfrow = c(1, 1))
} else {
  fit_kernel_main$message
}
```

### Flexible DML Level And Derivative Curves

`method = "dml_flex"` targets the flexible level effect from Section 3.4.1 of
the FDID notes:

$$
\theta(g) = E_X[\mu(g, X)].
$$

The implementation uses cross-fitting. For each fold, it fits nuisance models
for the outcome regression and the conditional density of `G` given `X`, builds
orthogonal pseudo-outcomes, then maps those pseudo-outcomes back to `G` using
`signal_map`. The derivative curve is obtained from the fitted level-curve map.
It is part of `dml_flex`, but it is more sensitive than the level curve.

Here all observations contribute to the cross-fitted pseudo-outcomes, but the
reported `theta(g)` and `theta_prime(g)` values are still evaluated only at
`eval_g`.

```{r mortality-dml-flex}
fit_flex_main <- fit_fdid(
  s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "dml_flex",
  learner = "linear",
  K = 3,
  S = 1,
  eval_g = eval_g_curve,
  signal_map = "spline",
  density_method = "residual_kde",
  boot = 3
)

if (!is_fit_error(fit_flex_main)) {
  par(mfrow = c(1, 2))
  plot(fit_flex_main, type = "curve", curve = "level",
       main = "DML flexible level curve")
  plot(fit_flex_main, type = "curve", curve = "derivative",
       main = "DML flexible derivative curve")
  par(mfrow = c(1, 1))
} else {
  fit_flex_main$message
}
```

The scalar `fit_flex_main$est$event` is a compatibility summary. The empirical
object to inspect is `fit_flex_main$curve_event`.

### Reporting Level Contrasts And Derivative Values

For applied reporting, the target is often not the whole curve but a specific
level comparison or derivative value. The helpers `fdid_contrast()` and
`fdid_derivative()` extract those quantities from the stored curve grid.

`fdid_contrast()` reports objects such as
`theta(g1) - theta(g0)` for `dml_flex` or `mu(g1) - mu(g0)` for `kernel`.
`fdid_derivative()` reports a fixed-`g` derivative value such as
`theta_prime(g0)`. When stored bootstrap or multiplier replicate curves are
available, the helpers use them. Otherwise they use conservative band-implied
intervals when possible, or clearly label pointwise endpoint-independent
intervals as approximate.

```{r mortality-reporting}
reporting_rows <- list()

if (!is_fit_error(fit_kernel_main)) {
  reporting_rows$kernel_contrast <- fdid_contrast(
    fit_kernel_main,
    g0 = min(fit_kernel_main$eval_g),
    g1 = max(fit_kernel_main$eval_g)
  )
  reporting_rows$kernel_derivative <- fdid_derivative(
    fit_kernel_main,
    g0 = stats::median(fit_kernel_main$eval_g)
  )
}

if (!is_fit_error(fit_flex_main)) {
  reporting_rows$flex_contrast <- fdid_contrast(
    fit_flex_main,
    g0 = min(fit_flex_main$eval_g),
    g1 = max(fit_flex_main$eval_g)
  )
  reporting_rows$flex_derivative <- fdid_derivative(
    fit_flex_main,
    g0 = stats::median(fit_flex_main$eval_g)
  )
}

if (length(reporting_rows) > 0) {
  reporting_table <- do.call(rbind, reporting_rows)
  knitr::kable(
    reporting_table[, c("target", "method", "g0", "g1", "estimate",
                        "std.error", "conf.low", "conf.high", "inference")],
    digits = 3,
    caption = "First-class reporting for continuous-G contrasts and derivative values"
  )
} else {
  "No curve fits are available for reporting."
}
```

## Scalar Average Derivative On Mortality

`method = "dml_incremental"` implements the separate Section 3.4.2 scalar
average-derivative target:

$$
E[\partial_g \mu(G, X)].
$$

This estimator is not obtained by averaging the derivative curve from
`dml_flex`. It uses its own orthogonal signal, which depends on both
`partial_g mu(G, X)` and `partial_g log s(G | X)`. Therefore the density method
is a substantive sensitivity choice.

```{r mortality-incremental}
fit_incremental_main <- list(
  residual_kde = fit_fdid(
    s_cont, tr_period = tr_period, ref_period = ref_period,
    method = "dml_incremental", learner = "linear", K = 3, S = 1,
    density_method = "residual_kde"
  ),
  location_scale = fit_fdid(
    s_cont, tr_period = tr_period, ref_period = ref_period,
    method = "dml_incremental", learner = "linear", K = 3, S = 1,
    density_method = "location_scale"
  )
)

incremental_table <- do.call(rbind, list(
  event_row(fit_incremental_main$residual_kde, "dml_incremental",
            "observed average derivative", learner = "linear",
            density_method = "residual_kde"),
  event_row(fit_incremental_main$location_scale, "dml_incremental",
            "observed average derivative", learner = "linear",
            density_method = "location_scale")
))

knitr::kable(incremental_table[, c("method", "learner", "density_method",
                                   "estimate", "std_error", "ci_lower",
                                   "ci_upper", "note")],
             digits = 4)
```

```{r mortality-incremental-summary}
if (!is_fit_error(fit_incremental_main$location_scale)) {
  summary(fit_incremental_main$location_scale)
}
```

## Sensitivity: HD Expansion, Learners, Density, Mapping

The main empirical story should not cross every option with every other option.
The sensitivity workflow varies one small set at a time:

- raw covariates versus high-dimensional basis expansion;
- learner family;
- density method for `dml_flex` and `dml_incremental`;
- signal map for `dml_flex`.

```{r learner-availability}
learner_table <- data.frame(
  learner = names(learner_package),
  package = unname(learner_package),
  available = vapply(names(learner_package), learner_available, logical(1))
)

main_render_learners <- c("linear")
heavy_render <- identical(tolower(Sys.getenv("FDID_TUTORIAL_HEAVY")), "true")
learner_table$included_in_render <- learner_table$available &
  (learner_table$learner %in% main_render_learners |
     (heavy_render & learner_table$learner %in%
        c("lasso", "elasticnet", "ranger", "grf", "nnet", "xgboost")))
learner_table$reason <- ifelse(
  learner_table$included_in_render,
  "included",
  ifelse(!learner_table$available, "optional package not installed",
         "set FDID_TUTORIAL_HEAVY=true for the full learner grid")
)

knitr::kable(learner_table, digits = 4)
```

```{r mortality-hd-objects}
s_cont_hd <- fdid_expand_covariates(
  s_cont,
  basis_type = "polynomial",
  poly_degree = 2,
  include_interactions = TRUE
)

hd_size_table <- data.frame(
  design = c("raw", "HD polynomial + interactions"),
  covariate_columns = c(
    length(grep("^x[0-9]+$", names(s_cont))),
    length(grep("^x[0-9]+$", names(s_cont_hd)))
  )
)
knitr::kable(hd_size_table)
```

### `dml_flex` Sensitivity

For `dml_flex`, the display should stay curve-first. The table below summarizes
the fitted curves, and the plot overlays the level curves from the included
specifications.

```{r mortality-dml-flex-sensitivity}
sens_eval_g <- eval_g_curve[round(seq(1, length(eval_g_curve), length.out = 7))]
sens_learners <- learner_table$learner[learner_table$included_in_render]
sens_learners <- intersect(sens_learners, names(learner_package))

fit_flex_sensitivity <- list()
for (design in c("raw", "HD")) {
  ss <- if (design == "raw") s_cont else s_cont_hd
  for (learner in sens_learners) {
    key <- paste(design, learner, sep = "_")
    fit_flex_sensitivity[[key]] <- fit_fdid(
      ss,
      tr_period = tr_period,
      ref_period = ref_period,
      method = "dml_flex",
      learner = learner,
      K = 3,
      S = 1,
      eval_g = sens_eval_g,
      signal_map = "spline",
      density_method = "location_scale",
      boot = 1
    )
  }
}

flex_sens_table <- do.call(rbind, lapply(names(fit_flex_sensitivity), function(key) {
  obj <- fit_flex_sensitivity[[key]]
  pieces <- strsplit(key, "_", fixed = TRUE)[[1]]
  design_i <- pieces[1]
  learner_i <- paste(pieces[-1], collapse = "_")
  if (is_fit_error(obj)) {
    return(data.frame(design = design_i, learner = learner_i,
                      mean_level = NA_real_, mean_derivative = NA_real_,
                      min_density = NA_real_, note = obj$message))
  }
  min_density_i <- if (!is.null(obj$dml_diagnostics$min_density)) {
    obj$dml_diagnostics$min_density
  } else {
    NA_real_
  }
  data.frame(
    design = design_i,
    learner = learner_i,
    mean_level = mean(obj$curve_event$theta_hat, na.rm = TRUE),
    mean_derivative = mean(obj$curve_event$delta_hat, na.rm = TRUE),
    min_density = min_density_i,
    note = ""
  )
}))

knitr::kable(flex_sens_table, digits = 4)
```

```{r mortality-dml-flex-sensitivity-plot}
flex_level_curves <- do.call(rbind, lapply(names(fit_flex_sensitivity), function(key) {
  obj <- fit_flex_sensitivity[[key]]
  pieces <- strsplit(key, "_", fixed = TRUE)[[1]]
  curve_row(obj, paste(pieces[1], paste(pieces[-1], collapse = "_"), sep = " / "),
            curve = "level")
}))

plot_curve_overlay(flex_level_curves, ylab = "theta_hat(g)",
                   main = "dml_flex level-curve sensitivity")
```

### `dml_incremental` Sensitivity

For `dml_incremental`, the main output is a scalar table. The two density
methods should be compared directly.

```{r mortality-dml-incremental-sensitivity}
incremental_learners <- intersect(sens_learners, c("linear", "gam", "ridge"))
inc_grid <- expand.grid(
  design = c("raw", "HD"),
  learner = incremental_learners,
  density_method = c("residual_kde", "location_scale"),
  stringsAsFactors = FALSE
)

fit_incremental_sensitivity <- vector("list", nrow(inc_grid))
for (ii in seq_len(nrow(inc_grid))) {
  ss <- if (inc_grid$design[ii] == "raw") s_cont else s_cont_hd
  fit_incremental_sensitivity[[ii]] <- fit_fdid(
    ss,
    tr_period = tr_period,
    ref_period = ref_period,
    method = "dml_incremental",
    learner = inc_grid$learner[ii],
    K = 3,
    S = 1,
    density_method = inc_grid$density_method[ii]
  )
}

inc_sens_table <- do.call(rbind, lapply(seq_len(nrow(inc_grid)), function(ii) {
  event_row(
    fit_incremental_sensitivity[[ii]],
    "dml_incremental",
    "observed average derivative",
    learner = inc_grid$learner[ii],
    design = inc_grid$design[ii],
    density_method = inc_grid$density_method[ii]
  )
}))

knitr::kable(inc_sens_table[, c("design", "learner", "density_method",
                                "estimate", "std_error", "ci_lower",
                                "ci_upper", "note")],
             digits = 4)
```

## Binary-G Companion

`method = "dml_binary"` is useful when the analyst intentionally dichotomizes
the factor, but it is not a continuous-$G$ estimand.

```{r mortality-binary-companion}
mortality_binary <- transform(
  mortality,
  G_binary = as.integer(lnpczupu > stats::median(lnpczupu, na.rm = TRUE))
)
s_binary <- fdid_prepare(
  mortality_binary,
  Y_label = "mortality",
  X_labels = covar,
  G_label = "G_binary",
  unit_label = "uniqueid",
  time_label = "year"
)

fit_binary_companion <- fit_fdid(
  s_binary,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "dml_binary",
  learner = "linear",
  K = 3,
  S = 1
)

knitr::kable(
  event_row(fit_binary_companion, "dml_binary", "binary AIPW/IRM effect",
            learner = "linear")[, c("method", "target", "estimate",
                                    "std_error", "ci_lower", "ci_upper", "note")],
  digits = 4
)
```

## Known-Truth Synthetic Sandbox

The synthetic examples below are diagnostics. They are not meant to replace the
oracle-signal validation scripts in `inst/`, but they make the failure modes
visible.

```{r synthetic-helpers}
make_panel <- function(delta_mu, n = 180, heteroskedastic = FALSE) {
  x1 <- stats::rnorm(n)
  x2 <- stats::runif(n, -1, 1)
  x3 <- stats::rbinom(n, 1, 0.45)
  x4 <- stats::rnorm(n)
  sd_g <- if (heteroskedastic) exp(-0.15 + 0.45 * x1) else rep(0.55, n)
  G <- 0.4 * x1 - 0.3 * x2 + 0.25 * x1 * x2 + stats::rnorm(n, sd = sd_g)
  mu <- delta_mu(G, x1, x2, x3, x4)
  unit_fe <- stats::rnorm(n)
  y1 <- unit_fe + stats::rnorm(n, sd = 0.35)
  dy <- mu + stats::rnorm(n, sd = 0.35)
  y2 <- y1 + dy
  y3 <- y1 + dy + stats::rnorm(n, sd = 0.20)
  data.frame(
    id = rep(seq_len(n), each = 3),
    time = rep(1:3, n),
    G = rep(G, each = 3),
    outcome = as.vector(rbind(y1, y2, y3)),
    x1 = rep(x1, each = 3),
    x2 = rep(x2, each = 3),
    x3 = rep(x3, each = 3),
    x4 = rep(x4, each = 3)
  )
}

prepare_synthetic <- function(d) {
  fdid_prepare(d, "outcome", c("x1", "x2", "x3", "x4"), "G", "id", "time")
}
```

### Example A: Smooth Nonlinear Dose-Response

This design has a smooth nonlinear curve. Scalar slopes are too compressed;
curve estimators show the shape.

```{r synthetic-a}
mu_a <- function(g, x1, x2, x3, x4) {
  1.2 * sin(g) - 0.35 * g^2 + 0.4 * x1 - 0.2 * x2 + 0.15 * x3
}
d_a <- make_panel(mu_a, n = 180)
s_a <- prepare_synthetic(d_a)
eval_a <- seq(stats::quantile(s_a$G, 0.10), stats::quantile(s_a$G, 0.90),
              length.out = 31)
u_a <- unique(d_a[, c("id", "x1", "x2", "x3", "x4")])
truth_a_level <- data.frame(
  G = eval_a,
  estimate = 1.2 * sin(eval_a) - 0.35 * eval_a^2 +
    mean(0.4 * u_a$x1 - 0.2 * u_a$x2 + 0.15 * u_a$x3)
)
truth_a_deriv <- data.frame(G = eval_a, estimate = 1.2 * cos(eval_a) - 0.7 * eval_a)

fit_a_kernel <- fit_fdid(s_a, tr_period = 2:3, ref_period = 1,
                         method = "kernel", eval_g = eval_a,
                         K_folds = 3, boot = 3)
fit_a_flex <- fit_fdid(s_a, tr_period = 2:3, ref_period = 1,
                       method = "dml_flex", learner = "gam",
                       K = 3, S = 1, eval_g = eval_a,
                       signal_map = "spline", density_method = "residual_kde",
                       boot = 3)

par(mfrow = c(1, 2))
plot_curve_overlay(
  rbind(curve_row(fit_a_kernel, "kernel", "level"),
        curve_row(fit_a_flex, "dml_flex", "level")),
  truth = truth_a_level, ylab = "level", main = "Smooth level"
)
plot_curve_overlay(
  rbind(curve_row(fit_a_kernel, "kernel", "derivative"),
        curve_row(fit_a_flex, "dml_flex", "derivative")),
  truth = truth_a_deriv, ylab = "derivative", main = "Smooth derivative"
)
par(mfrow = c(1, 1))
```

### Example B: Nonlinear Confounding And HD Expansion

This design lets the effect of `G` interact with nonlinear functions of `X`.
The comparison focuses on `dml_flex` with raw versus expanded covariates.

```{r synthetic-b}
mu_b <- function(g, x1, x2, x3, x4) {
  sin(g) + 0.45 * g * x1^2 - 0.25 * g * x2 +
    0.4 * x1^2 - 0.3 * x2 + 0.2 * x3 * x4
}
d_b <- make_panel(mu_b, n = 180)
s_b_raw <- prepare_synthetic(d_b)
s_b_hd <- fdid_expand_covariates(s_b_raw, basis_type = "polynomial",
                                 poly_degree = 2, include_interactions = TRUE)
eval_b <- seq(stats::quantile(s_b_raw$G, 0.15), stats::quantile(s_b_raw$G, 0.85),
              length.out = 21)
u_b <- unique(d_b[, c("id", "x1", "x2", "x3", "x4")])
truth_b <- data.frame(
  G = eval_b,
  estimate = sin(eval_b) +
    0.45 * eval_b * mean(u_b$x1^2) -
    0.25 * eval_b * mean(u_b$x2) +
    mean(0.4 * u_b$x1^2 - 0.3 * u_b$x2 + 0.2 * u_b$x3 * u_b$x4)
)

fit_b_raw <- fit_fdid(s_b_raw, tr_period = 2:3, ref_period = 1,
                      method = "dml_flex", learner = "gam", K = 3, S = 1,
                      eval_g = eval_b, signal_map = "spline",
                      density_method = "location_scale", boot = 3)
fit_b_hd <- fit_fdid(s_b_hd, tr_period = 2:3, ref_period = 1,
                     method = "dml_flex", learner = "ridge", K = 3, S = 1,
                     eval_g = eval_b, signal_map = "spline",
                     density_method = "location_scale", boot = 3)

plot_curve_overlay(
  rbind(curve_row(fit_b_raw, "raw X / gam", "level"),
        curve_row(fit_b_hd, "HD X / ridge", "level")),
  truth = truth_b, ylab = "theta(g)", main = "HD expansion under nonlinear confounding"
)
```

### Example C: Heteroskedastic `G | X`

This design makes the conditional density of `G` depend on `X`. The scalar
average derivative is compared across density methods.

```{r synthetic-c}
mu_c <- function(g, x1, x2, x3, x4) {
  1.2 * g - 0.4 * g^2 + 0.5 * g * x1 + 0.25 * x2^2
}
d_c <- make_panel(mu_c, n = 220, heteroskedastic = TRUE)
s_c <- prepare_synthetic(d_c)
truth_c <- with(unique(d_c[, c("id", "G", "x1")]),
                mean(1.2 - 0.8 * G + 0.5 * x1))

fit_c_resid <- fit_fdid(s_c, tr_period = 2:3, ref_period = 1,
                        method = "dml_incremental", learner = "gam",
                        K = 3, S = 1, density_method = "residual_kde")
fit_c_loc <- fit_fdid(s_c, tr_period = 2:3, ref_period = 1,
                      method = "dml_incremental", learner = "gam",
                      K = 3, S = 1, density_method = "location_scale")

synthetic_c_table <- rbind(
  event_row(fit_c_resid, "dml_incremental", "truth comparison",
            learner = "gam", density_method = "residual_kde"),
  event_row(fit_c_loc, "dml_incremental", "truth comparison",
            learner = "gam", density_method = "location_scale")
)
synthetic_c_table$truth <- truth_c
synthetic_c_table$error <- synthetic_c_table$estimate - truth_c
knitr::kable(synthetic_c_table[, c("learner", "density_method", "estimate",
                                   "std_error", "truth", "error", "note")],
             digits = 4)
```

### Example D: Kinked Outcome Function

Derivative curves require smoothness. A kink can leave the level curve
informative while making derivative plots hard to read near the kink.

```{r synthetic-d}
mu_d <- function(g, x1, x2, x3, x4) {
  abs(g - 0.2) + 0.3 * x1 - 0.25 * x2
}
d_d <- make_panel(mu_d, n = 180)
s_d <- prepare_synthetic(d_d)
eval_d <- seq(stats::quantile(s_d$G, 0.10), stats::quantile(s_d$G, 0.90),
              length.out = 31)
truth_d_level <- data.frame(
  G = eval_d,
  estimate = abs(eval_d - 0.2) + mean(0.3 * s_d$x1 - 0.25 * s_d$x2)
)
truth_d_deriv <- data.frame(G = eval_d, estimate = ifelse(eval_d < 0.2, -1, 1))

fit_d_kernel <- fit_fdid(s_d, tr_period = 2:3, ref_period = 1,
                         method = "kernel", eval_g = eval_d,
                         K_folds = 3, boot = 3)
fit_d_flex <- fit_fdid(s_d, tr_period = 2:3, ref_period = 1,
                       method = "dml_flex", learner = "gam",
                       K = 3, S = 1, eval_g = eval_d,
                       signal_map = "spline", density_method = "residual_kde",
                       boot = 3)

par(mfrow = c(1, 2))
plot_curve_overlay(
  rbind(curve_row(fit_d_kernel, "kernel", "level"),
        curve_row(fit_d_flex, "dml_flex", "level")),
  truth = truth_d_level, ylab = "level", main = "Kinked level"
)
plot_curve_overlay(
  rbind(curve_row(fit_d_kernel, "kernel", "derivative"),
        curve_row(fit_d_flex, "dml_flex", "derivative")),
  truth = truth_d_deriv, ylab = "derivative", main = "Kinked derivative"
)
par(mfrow = c(1, 1))
```

### Synthetic Learner And HD Grid

The grid below uses the nonlinear-confounding synthetic design because the
truth is known and HD expansion should help. It compares all installed learner
families in a small render-safe design.

```{r synthetic-all-learners}
synth_learners <- if (heavy_render) {
  learner_table$learner[learner_table$available]
} else {
  intersect(c("linear", "gam", "ridge", "ranger"),
            learner_table$learner[learner_table$available])
}
synth_eval <- eval_b[round(seq(1, length(eval_b), length.out = 9))]

fit_synth_grid <- list()
for (design in c("raw", "HD")) {
  ss <- if (design == "raw") s_b_raw else s_b_hd
  for (learner in synth_learners) {
    key <- paste(design, learner, sep = "_")
    fit_synth_grid[[key]] <- fit_fdid(
      ss,
      tr_period = 2:3,
      ref_period = 1,
      method = "dml_flex",
      learner = learner,
      K = 2,
      S = 1,
      eval_g = synth_eval,
      signal_map = "spline",
      density_method = "location_scale",
      boot = 1
    )
  }
}

truth_synth <- approx(truth_b$G, truth_b$estimate, xout = synth_eval,
                      rule = 2)$y
synth_grid_table <- do.call(rbind, lapply(names(fit_synth_grid), function(key) {
  obj <- fit_synth_grid[[key]]
  pieces <- strsplit(key, "_", fixed = TRUE)[[1]]
  design_i <- pieces[1]
  learner_i <- paste(pieces[-1], collapse = "_")
  if (is_fit_error(obj)) {
    return(data.frame(design = design_i, learner = learner_i,
                      mae_level = NA_real_, max_error = NA_real_,
                      note = obj$message))
  }
  err <- obj$curve_event$theta_hat - truth_synth
  data.frame(design = design_i, learner = learner_i,
             mae_level = mean(abs(err), na.rm = TRUE),
             max_error = max(abs(err), na.rm = TRUE),
             note = "")
}))

knitr::kable(synth_grid_table[order(synth_grid_table$design,
                                    synth_grid_table$mae_level), ],
             digits = 4)
```

## Method Choice Guide

| Situation | Start with | Report |
|---|---|---|
| Low-dimensional smooth relationship | `method = "kernel"` | level and derivative curves |
| Need a scalar benchmark | `method = "dml_plr"` | scalar slope |
| Need a nonlinear adjusted curve | `method = "dml_flex"` | level curve; derivative as sensitivity |
| Need the observed average derivative | `method = "dml_incremental"` | scalar estimate across density methods |
| `G | X` may be heteroskedastic | `density_method = "location_scale"` | density sensitivity |
| Covariates need richer adjustment | `fdid_expand_covariates()` | learner and HD sensitivity |
| Density methods disagree | do not report one estimate alone | support and density diagnostics |

The practical diagnostic message is simple: use scalar estimators for scalar
questions, use `kernel` and `dml_flex` for curves, and treat derivative targets
as more fragile than level targets.
