# DML Estimators for Continuous G {#sec-dml-continuous}

```{r dml-setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 7,
  fig.height = 4.8
)
set.seed(20260531)

fit_fdid <- function(...) {
  tryCatch(
    fdid(...),
    error = function(e) {
      structure(list(message = conditionMessage(e)), class = "fdid_fit_error")
    }
  )
}

is_fit_error <- function(x) inherits(x, "fdid_fit_error")

event_row <- function(obj, label, target, learner = NA_character_,
                      density_method = NA_character_,
                      signal_map = NA_character_) {
  if (is_fit_error(obj)) {
    return(data.frame(
      fit = label, target = target, learner = learner,
      density_method = density_method, signal_map = signal_map,
      Estimate = NA_real_, Std.Error = NA_real_,
      CI_Lower = NA_real_, CI_Upper = NA_real_,
      note = obj$message,
      check.names = FALSE
    ))
  }
  ev <- obj$est$event
  data.frame(
    fit = label, target = target, learner = learner,
    density_method = density_method, signal_map = signal_map,
    Estimate = as.numeric(ev$Estimate[1]),
    Std.Error = as.numeric(ev$Std.Error[1]),
    CI_Lower = as.numeric(ev$CI_Lower[1]),
    CI_Upper = as.numeric(ev$CI_Upper[1]),
    note = "",
    check.names = FALSE
  )
}

metadata_table <- function(obj) {
  if (is_fit_error(obj) || is.null(obj$dml_metadata)) return(data.frame())
  vals <- vapply(obj$dml_metadata, function(z) {
    if (length(z) == 0L || is.null(z)) return(NA_character_)
    paste(as.character(z), collapse = ", ")
  }, character(1L))
  data.frame(field = names(vals), value = unname(vals), row.names = NULL)
}
```

This chapter gives a focused workflow for the double/debiased machine learning
(DML) estimators in `fdid`. It follows the teaching structure used in the
`interflex` DML tutorial and @liu2025practical: first map the target, then
explain nuisance learning and orthogonal signals, then separate scalar,
curve, and incremental estimands.

The most important rule is that the DML methods are not interchangeable labels
for one object. They target different quantities.

## Target Map

```{r dml-target-map}
dml_targets <- data.frame(
  method = c("dml_plr", "dml_flex", "dml_incremental", "dml_binary"),
  target = c(
    "Partially linear scalar slope",
    "Flexible level curve theta(g) and derivative theta'(g)",
    "Observed-population average derivative E[partial_g mu(G, X)]",
    "Binary-G AIPW/IRM contrast"
  ),
  main_output = c(
    "summary table",
    "curve plots and reporting helpers",
    "summary table",
    "summary table"
  ),
  main_caveat = c(
    "Scalar PLR target only",
    "Derivative and curve bands are more fragile than levels",
    "Depends on the conditional-density score",
    "Requires G in {0, 1}; not a continuous-G estimator"
  )
)

knitr::kable(dml_targets)
```

All DML estimators work with the FDID first-difference outcome. Conceptually,
the event-period target is built from

$$
\Delta Y_i = \bar Y_{i,\mathrm{event}} - Y_{i,\mathrm{ref}}.
$$

DML then proceeds in three steps:

1. Learn nuisance functions using cross-fitting.
2. Construct an orthogonal signal for the target.
3. Aggregate or map the signal to the target scale: a scalar slope, a smooth
   curve in `G`, or an observed-population average derivative.

## Data Setup

```{r dml-library-data}
library(fdid)
data(fdid)

mortality$uniqueid <- paste(
  as.character(mortality$provid),
  as.character(mortality$countyid),
  sep = "-"
)

covar <- c("avggrain", "nograin", "urban", "dis_bj", "dis_pc",
           "rice", "minority", "edu", "lnpop")
tr_period <- 1958:1961
ref_period <- 1957

s_cont <- fdid_prepare(
  data = transform(mortality, G_cont = lnpczupu),
  Y_label = "mortality",
  X_labels = covar,
  G_label = "G_cont",
  unit_label = "uniqueid",
  time_label = "year"
)

g_positive <- s_cont$G[s_cont$G > 0]
eval_g_dml <- as.numeric(stats::quantile(
  g_positive,
  probs = c(0.20, 0.35, 0.50, 0.65, 0.80),
  na.rm = TRUE
))
```

As in Chapter @sec-kernel-continuous, `eval_g` is an estimation grid for
`dml_flex`. The scalar DML methods ignore it because they do not estimate a
curve.

## Scalar PLR Benchmark

`method = "dml_plr"` estimates a partially linear scalar slope. It is a useful
benchmark when the question is whether the first-difference outcome has one
average adjusted relationship with continuous `G`.

```{r dml-plr-fit}
fit_plr_score <- fit_fdid(
  s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "dml_plr",
  learner = "linear",
  K = 3,
  S = 1,
  dml_inference = "score"
)

fit_plr_multiplier <- fit_fdid(
  s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "dml_plr",
  learner = "linear",
  K = 3,
  S = 1,
  dml_inference = "score_multiplier",
  dml_boot = 99
)

plr_table <- rbind(
  event_row(fit_plr_score, "PLR score", "partially linear slope",
            learner = "linear"),
  event_row(fit_plr_multiplier, "PLR score multiplier",
            "partially linear slope", learner = "linear")
)

knitr::kable(plr_table[, c("fit", "target", "learner", "Estimate",
                           "Std.Error", "CI_Lower", "CI_Upper", "note")],
             digits = 3)
```

The default scalar DML interval is analytical score inference. The
`score_multiplier` option changes the critical value using a scalar multiplier
process based on stored score contributions. It is not a nonparametric refit
bootstrap.

```{r dml-plr-metadata}
if (!is_fit_error(fit_plr_multiplier)) {
  knitr::kable(metadata_table(fit_plr_multiplier))
}
```

## Flexible DML Curves

`method = "dml_flex"` targets the flexible level curve

$$
\theta(g) = E_X[\mu(g, X)]
$$

and its derivative with respect to `G`. The implementation cross-fits nuisance
models, builds orthogonal pseudo-outcomes, and maps those pseudo-outcomes back
to the `eval_g` grid using `signal_map`.

Two option families are easy to confuse:

- `density_method` estimates the conditional density $f_{G \mid X}(g \mid x)$
  needed by the orthogonal continuous-treatment score. It is used by
  `dml_flex` and `dml_incremental`.
- `signal_map` maps the cross-fitted `dml_flex` signal back to the `eval_g`
  grid. It is used by `dml_flex` only.

`signal_map = "blp_spline"` is therefore not a fourth density algorithm. It is
a B-spline best linear projection mapping path that also stores finite-grid
covariance matrices for level and derivative curves.

### Conditional-density options

The package currently exposes three conditional-density approximations.

| `density_method` | Algorithm | Practical reading |
|---|---|---|
| `"residual_kde"` | Fit $E[G \mid X]$, compute residuals, and apply a univariate KDE to $G - \hat m(X)$ | fast default; works best when `G | X` is close to a location-shift model |
| `"location_scale"` | Fit $E[G \mid X]$, fit a residual-scale model, standardize residuals, and KDE the standardized residual | allows heteroskedastic `G | X`; often preferable in heteroskedastic designs |
| `"local_kde"` | Product Gaussian kernel over scaled covariates and a Gaussian KDE over `G` | most nonparametric of the three; sensitive to covariate dimension |

The density estimate enters the DML score, not just a plot. For this reason,
`density_method` should be treated as a sensitivity dimension, especially for
`dml_incremental`, where the target is an observed-population average
derivative and the conditional-density score is central.

```{r dml-flex-fit}
dml_curve_draws <- if (identical(tolower(Sys.getenv("FDID_TUTORIAL_HEAVY")),
                                 "true")) 199L else 49L

fit_flex_spline <- fit_fdid(
  s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "dml_flex",
  learner = "linear",
  K = 3,
  S = 1,
  eval_g = eval_g_dml,
  signal_map = "spline",
  density_method = "residual_kde",
  boot = dml_curve_draws
)

fit_flex_blp <- fit_fdid(
  s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "dml_flex",
  learner = "linear",
  K = 3,
  S = 1,
  eval_g = eval_g_dml,
  signal_map = "blp_spline",
  density_method = "location_scale",
  boot = dml_curve_draws
)

flex_table <- rbind(
  event_row(fit_flex_spline, "dml_flex spline", "average level summary",
            learner = "linear", density_method = "residual_kde",
            signal_map = "spline"),
  event_row(fit_flex_blp, "dml_flex BLP spline", "average level summary",
            learner = "linear", density_method = "location_scale",
            signal_map = "blp_spline")
)

knitr::kable(flex_table[, c("fit", "signal_map", "density_method",
                            "Estimate", "Std.Error", "CI_Lower", "CI_Upper",
                            "note")],
             digits = 3)
```

```{r dml-flex-plots, fig.width=9, fig.height=4.4}
if (!is_fit_error(fit_flex_spline)) {
  old_par <- par(mfrow = c(1, 2))
  plot(fit_flex_spline, type = "curve", curve = "level",
       interval = "both", Xdistr = "rug",
       main = "DML flexible level curve")
  plot(fit_flex_spline, type = "curve", curve = "derivative",
       interval = "both", Xdistr = "rug",
       main = "DML flexible derivative curve")
  par(old_par)
}
```

The scalar event row for `dml_flex` is a compatibility summary. The curve fields
are the primary output.

```{r dml-flex-curve-table}
if (!is_fit_error(fit_flex_spline)) {
  flex_curve <- data.frame(
    G = fit_flex_spline$eval_g,
    theta_hat = fit_flex_spline$curve_event$theta_hat,
    se_theta = fit_flex_spline$curve_event$se_theta,
    delta_hat = fit_flex_spline$curve_event$delta_hat,
    se_delta = fit_flex_spline$curve_event$se_delta
  )
  knitr::kable(flex_curve, digits = 3)
}
```

## DML Curve Inference

The DML curve inference labels are deliberately method-specific.

- Pointwise intervals for `dml_flex` are second-stage mapping intervals.
- With `signal_map = "spline"`, the simultaneous band is a practical residual
  multiplier band over cross-fitted signal-mapping residuals.
- With `signal_map = "blp_spline"`, the object stores spline-basis covariance
  matrices and uses a Gaussian max-$t$ critical value for finite-grid bands
  when the covariance is usable.
- Neither path is a full nonparametric DML refit bootstrap over all nuisance
  stages.

```{r dml-flex-inference-metadata}
if (!is_fit_error(fit_flex_blp)) {
  inference_rows <- data.frame(
    field = c("signal_map", "density_method", "band_method",
              "theta_vcov_available", "delta_vcov_available",
              "gaussian_crit_theta", "simultaneous_crit_theta"),
    value = c(
      fit_flex_blp$signal_map,
      fit_flex_blp$density_method,
      fit_flex_blp$curve_event$simultaneous_band_method,
      !is.null(fit_flex_blp$curve_event$theta_vcov),
      !is.null(fit_flex_blp$curve_event$delta_vcov),
      round(fit_flex_blp$curve_event$gaussian_crit_theta, 3),
      round(fit_flex_blp$curve_event$simultaneous_crit_theta, 3)
    )
  )
  knitr::kable(inference_rows)
}
```

```{r dml-blp-band-plot, fig.width=7, fig.height=4.5}
if (!is_fit_error(fit_flex_blp)) {
  plot(fit_flex_blp, type = "curve", curve = "level",
       interval = "both", Xdistr = "density",
       main = "BLP-spline Gaussian max-t band")
}
```

## Reporting Curve Quantities

Use the reporting helpers when the estimand is a specific level contrast or
derivative value. For BLP-spline curves, `fdid_contrast()` can use the stored
curve covariance matrix, so exact-grid contrasts reduce to the familiar
`V22 + V11 - 2 * V12` identity.

```{r dml-reporting}
if (!is_fit_error(fit_flex_blp)) {
  dml_report <- rbind(
    fdid_contrast(fit_flex_blp, g0 = min(eval_g_dml), g1 = max(eval_g_dml)),
    fdid_derivative(fit_flex_blp, g0 = stats::median(eval_g_dml))
  )

  knitr::kable(
    dml_report[, c("target", "method", "g0", "g1", "estimate",
                   "std.error", "conf.low", "conf.high", "inference")],
    digits = 3
  )
}
```

## Incremental Average Derivative

`method = "dml_incremental"` is not the average of the `dml_flex` derivative
curve. It targets the observed-population average derivative directly, using an
orthogonal score that includes the conditional density score for `G | X`.
Therefore `density_method` is a substantive sensitivity choice.

```{r dml-incremental-fit}
fit_inc_residual <- fit_fdid(
  s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "dml_incremental",
  learner = "linear",
  K = 3,
  S = 1,
  density_method = "residual_kde"
)

fit_inc_location <- fit_fdid(
  s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "dml_incremental",
  learner = "linear",
  K = 3,
  S = 1,
  density_method = "location_scale"
)

fit_inc_local <- fit_fdid(
  s_cont,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "dml_incremental",
  learner = "linear",
  K = 3,
  S = 1,
  density_method = "local_kde"
)

incremental_table <- rbind(
  event_row(fit_inc_residual, "incremental residual KDE",
            "observed average derivative", learner = "linear",
            density_method = "residual_kde"),
  event_row(fit_inc_location, "incremental location-scale",
            "observed average derivative", learner = "linear",
            density_method = "location_scale"),
  event_row(fit_inc_local, "incremental local KDE",
            "observed average derivative", learner = "linear",
            density_method = "local_kde")
)

knitr::kable(incremental_table[, c("fit", "density_method", "Estimate",
                                   "Std.Error", "CI_Lower", "CI_Upper",
                                   "note")],
             digits = 3)
```

## Learner And Design Sensitivity

The DML estimators support several nuisance learners. Optional learner packages
are loaded only when requested, so a tutorial should check availability before
running a large grid.

```{r dml-learner-availability}
learner_package <- c(
  linear = NA_character_,
  gam = "mgcv",
  lasso = "glmnet",
  ridge = "glmnet",
  elasticnet = "glmnet",
  ranger = "ranger",
  grf = "grf",
  nnet = "nnet",
  xgboost = "xgboost"
)

learner_available <- function(learner) {
  pkg <- learner_package[[learner]]
  is.na(pkg) || requireNamespace(pkg, quietly = TRUE)
}

learner_table <- data.frame(
  learner = names(learner_package),
  package = unname(learner_package),
  available = vapply(names(learner_package), learner_available, logical(1)),
  stringsAsFactors = FALSE
)

knitr::kable(learner_table)
```

High-dimensional expansions are another sensitivity dimension. The example
below creates a polynomial and interaction basis without running the full
learner grid.

```{r dml-hd-design}
s_cont_hd <- fdid_expand_covariates(
  s_cont,
  basis_type = "polynomial",
  poly_degree = 2,
  include_interactions = TRUE
)

hd_table <- data.frame(
  design = c("raw covariates", "polynomial plus interactions"),
  covariate_columns = c(
    length(grep("^x[0-9]+$", names(s_cont))),
    length(grep("^x[0-9]+$", names(s_cont_hd)))
  )
)

knitr::kable(hd_table)
```

For a full analysis, vary one dimension at a time: learner, expanded design,
`density_method`, then `signal_map`. The rendered tutorial keeps the grid
small; set `FDID_TUTORIAL_HEAVY=true` before rendering if you want larger
replication counts or optional learner grids.

## DML Checklist

Before reporting a DML result, check:

- whether the target is scalar, curve, or incremental;
- whether `G` is truly continuous or binary;
- which nuisance learner was used and whether optional packages were loaded;
- which conditional-density method was used for `dml_flex` or
  `dml_incremental`;
- whether the interval is analytical score, scalar score multiplier,
  practical signal-residual multiplier, or BLP-spline Gaussian max-$t$;
- whether curve conclusions rely on sparse support or a fragile derivative.
